### 2020-09-05 chatbot_classification/aaai_2021/paper.tex
update experimental results and analysis

\def\year{2021}\relax
%File: formatting-instructions-latex-2021.tex
%release 2021.1
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai21}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
\pdfinfo{
  /Title (AAAI Press Formatting Instructions for Authors Using LaTeX -- A Guide)
  /Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen, Francisco Cruz, Marc Pujol-Gonzalez)
  /TemplateVersion (2021.1)
}

\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfigure}
\usepackage{spconf}

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

\def\x{{\mathbf x}}
\def\L{{\cal L}}

\title{     
  %SFC: Few-Shot Text Classification via Similarity Fused with Classification System
  Joint training of classification model and similarity model for low-resource
  text classification in chatbot
}

\author{
  %Authors
  % All authors must be in the same font size and format.
  Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
  AAAI Style Contributions by Pater Patel Schneider,
  Sunil Issar,  \\
  J. Scott Penberthy,
  George Ferguson,
  Hans Guesgen,
  Francisco Cruz,
  Marc Pujol-Gonzalez
  \\
}
\affiliations{
  %Afiliations

  \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
  %If you have multiple authors and multiple affiliations
  % use superscripts in text and roman font to identify them.
  %For example,

  % Sunil Issar, \textsuperscript{\rm 2}
  % J. Scott Penberthy, \textsuperscript{\rm 3}
  % George Ferguson,\textsuperscript{\rm 4}
  % Hans Guesgen, \textsuperscript{\rm 5}.
  % Note that the comma should be placed BEFORE the superscript for optimum readability

  2275 East Bayshore Road, Suite 160\\
  Palo Alto, California 94303\\
  % email address must be in roman text type, not monospace or sans serif
  publications21@aaai.org

  % See more examples next
}

\begin{document}
  \maketitle

  \begin{abstract}
    Task-specific  chatbot  systems has gained many important applications, such
    as  Siri, Google home, customer service systems to allievaite human's labor.
    In  this  kind  of  systems, one of the core problems is the accurate detection of
    intention  of  a  user's input. It is then mapped into a predefined chatting
    logic  graph,  and the chatbot can return reasonable answers easily. This is
    the  main  difference from a free chatbot. The intention detection, assuming
    only  one  intention  in  a  sentence,  can  be  modeled  as  a  short  text
    classification  problem.  In  practice,  the  number  of  intentions  can be
    hundreads of thousands, as opposed to quite limited in the tradiational text
    classification  problem,  and  the  labeled  data  is not trivial to acquire
    plenty,  generally  10  to 20 per intention. Then, how to train an efficient
    model  in such a resource-intensive practical application in a new domain is
    critical.  In  this  work, we propose a novel model, called similairty model
    fused with classification Model (SFC), which combines a classification model
    and  a  similairty  model,  and  borrows the power of multi-task training as
    well.  We  conduct extensive experiments on 4 public and 1 private datasets.
    The  experimental  results  show  that our new model outperforms very strong
    baselines, BERT, RoBERTa and ALBERT based pretrained models by 2 percentages
    in average accuracy.

  \end{abstract}

  \section{Introduction}
  \label{sec:intro}

  Task-specific  conversational  chatbot  \cite{wen2016network}  has been applied
  into  many practical products. For example, it can share the working burden of
  human  customer  service  agents  responsible  for  customers' questions about
  certain  products  or services. Another example is the speech-based control in
  devices  like  intelligent  speakers, such as Siri, Google home, Baidu Xiaodu,
  and smart home equipments. Regardless of whether these conversations belong to
  single-turn or multi-turn, an essential technique underlying the chatbot is to
  identify  the  real  intention  behind  a  user's  question or reply. Then the
  chatbot  is caple of conducting expected actions, text reply or accomplish some
  operations.

  It  is  a  challenging task due to some inherent properties. First,
  a customer's  utterance  within  a conversation is usually quite short. Since it
  does  not  contain  enough  information,  short  text \cite{song2014short} are
  thought  to  be  more  ambiguous  compared to long texts such as paragraphs or
  documents,  poseing  a  great challenge \cite{chen2019deep} for classification
  task   \cite{phan2008learning,yan2009dynamic,hua2015short}.   Second,  in  the
  initial  stage  of  building  a  chatbot  , it is often rather hard to collect
  sufficient  labeled  data  to  train  a good model. Because developers have to
  examine tons of real system logs for typical user's utterances, and then label
  them  with expected intentions. Therefore, the key to building a task-specific
  chatbot  with  high  performance  becomes  solving the challenge of short-text
  classification   \cite{sriram2010short}   problem   under   few-shot   setting
  \cite{yu2018diverse}.

  Generally  there  are  two  kinds  of  approaches  to  address  this  problem,
  classification model that directly maps an input into a most likely label, and
  similarity  model  that searches a label whose associated data is most similar to
  an input.

  The   first  one,  \emph{text  classification},  includes  classic  long  text
  classification  and  short  text  classification  in  this  scenario.  Besides
  traditional  machine  learning  models  like  SVM,  boosting  tree,  many work
  \cite{wen2016network}  choose  neural  network  such  as  convolutional neural
  networks (CNNs) \cite{kim2014convolutional,zhang2015character,conneau2016very}
  and       long       short       term       memory       networks      (LSTMs)
  \cite{mousa2017contextual,liu2016recurrent}   to   accomplish   the   task  of
  extracting  semantic  feature from limited amount of words. Their common model
  structure  is  adding  a  softmax classifier as the final layer and mapping to
  labels.  Another  two  interesting lines of work, label-word joint models, and
  joint  NER  and  classification  are  discussed  in  the related work section.
  Afterwards,   pre-trained   language   models   on   large  corpus  like  BERT
  \cite{devlin2018bert}  and  RoBERTa \cite{liu2019roberta} has been proven more
  powerful  in  solving  many  NLP  tasks  including  short-text  classification
  \cite{madabushi2020cost}.      Especially      for      few-shot     scenarios
  \cite{yu2018diverse},     pre-trained     model    based    on    transformers
  \cite{vaswani2017attention}, shown in Fig. \ref{Fig.1.sub.1}, tends to do more
  help to reducing the negative effects brought by scarcity of training data.

  The  second  one,  \emph{similarity model} is also called sentence-pair model.
  Its  motivation  originates  from  the idea of Information Retrival (IR) based
  chatbot   \cite{jafarpour2010filter,   leuski2011npceditor}.   Having   a  Q-A
  (Question-Answer)  pairs dataset and user query Q, the IR based conversational
  system will look up in the Q-A dataset for the pair (Q', A') that best matches
  query  Q  through  semantic  analysis  and  returns  A'  as  the  answer  to Q
  \cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large
  corpus  of  semantic  similarity  identification task can be applied for being
  used  as  a  tool  to  identify  the class with highest semantic similarity to
  customer's    query.    A    common    model    structure    of    pre-trained
  \cite{devlin2018bert} sentence-pair model is based on multiple cross-attention
  mechanism  \cite{barkan2020scalable},  shown in Fig. \ref{Fig.1.sub.2}. Due to
  the   fact   that   many   experiment   results   have   shown   that  RoBERTa
  \cite{liu2019roberta} is an improved version of BERT, and has achieved amazing
  results  on  both  text  classification and sentence-pairs semantic similarity
  tasks, we choose RoBERTa as an important baseline approach in this paper.

  Despite  the  success  of  these two kinds of approaches, they still have some
  limitations   in   task-specific  chatbot  scenario.  As  for  the  \emph{text
  classification  model},  it  is  quite hard to use data augmentation method to
  solve  the  data insufficiency challenge since the it is usually unfeasible to
  get  large amount of domain-specific data when facing a new task. With respect
  to  the  \emph{similarity model}, though we can obtain a large amount of data
  from semantically  duplicate  sentence  pair identification domain for transfer
  learning  \cite{sun2019fine},  it  is  still  quite  hard  to  perform well on
  a new task-specific dataset,  since their objective losses are totally different.

  The  above limitation motivates us to propose a joint system that is capble of
  taking  advantagges  from  both  classification model and similarity model. We
  calle  it  SFC,  short  for  similairty model fused with classification Model,
  shown  in  Fig.  \ref{fig:ver.1.3}.  To  better train SFC, we further bring in
  multi-task   learning   \cite{caruana1993multitask,collobert2008unified}.  Our
  basic  idea  cames  from  two  stages. In the first stage, we use an auxiliary
  model  to  select top-K most possible labels for a input. This model can be an
  elasticsearch  \cite{divya2013elasticsearch}  or  a  text classification model
  trained on current task-specific chatbot data. In the second stage, we build a
  classification  model  whose  main modules are similarity models. Then, this
  structure  derives two goals to train towards, the classification loss for our
  application,  and the similarity loss for the main modules. When the two stages
  are  independently  optimized,  we call \emph{two-stage SFC}. We further find,
  the  the  quality  of  output  from  the  first  stage  might  limit the final
  performance  of  system,  since  the  candidate  labels for each user query is
  always  fixed  in  the  whole  multi-task  training  process. This observation
  motivates  us  to  further  improve  the  two-stage SFC into a joint training,
  called  \emph{joint  SFC}. In this way, the sentence pairs associated with the
  top-K  classes  also  become dynamic, and the text classification model in the
  first  stage  will also be further optimized to provide better top-K result by
  the final training loss.

  We summarize our contributions as follow:
  \begin{enumerate}
    \item We propose a novel joint system to makes full utilization of
    the advantages from both text classification model and similarity model.

    \item To  better  train such a system, we bring in multi-task training to obtain
    expected performance.

    \item Experiment  results  from 4 public and 1 private short-text classification
    datasets,  show  that  our  proposed  SFC joint system can achieve significant
    improvements over strong baselines, especially in the low resource settings.
  \end{enumerate}

  \begin{figure*}[t]
    \begin{centering}
      \includegraphics[scale=0.72]{picture/picture4} 
      \par
    \end{centering}
    \caption{
      \textbf{Network   Structure   of  joint  SFC:}  a  joint  model  of
      sentence-pair   model   and   text   classification   model  organized  in
      joint  structure by completing different tasks at different network
      layers.
    }
    \label{fig:framework}
  \end{figure*}

  \section{Related work}

  Another  common  approach  is  text  classification  neural  network  based on
  label-word     joint     embedding,    such    as    LEAM \cite{wang2018joint},
  MTLE \cite{zhang2017multi}   and   EXAM \cite{du2019explicit}.   This   approach
  involves  label  embedding when constructing the text-sequence representation.
  Take  LEAM  as  an  example,  the  text classification model is implemented by
  jointly    embedding    the    word    and    label   in   the   same   latent
  space \cite{wang2018joint}. Afterwards, the text representation are constructed
  based  on  text-label  compatibility through attention mechanism. In this way,
  additional  sources  of  semantic  information  from class labels can be fully
  leveraged, which is quite helpful especially under few-shot scenario.

  As for \emph{label-word joint
  embedding  approach},  it  poses  noticeable requirements on label information.
  However,  in  our  cases,  since  we have large amount of class labels without
  being well-defined for question answering chatbot, the only thing we can do is
  to  choose  one  of the user utterance as the class label to represent the key
  information  of  the  whole class. In this way, the label will become not only
  prolix  but  also  biased,  which  will  bring  some  negative effect to label
  embedding.

  {Add joint NER and classification problem here.}

  \section{Methodology}
  In  this  section,  we  will  describe our proposed system of similarity model fused with
  classification  model,  named  as  SFC, in detail. We come up with 2 different
  approaches  for  implementation  of SFC system, in which joint SFC is a
  further improvement of two-stage SFC for better performance.

  \subsection{two-stage SFC with multi-task learning}

  The network   structure   of   two-stage   SFC  is  composed  with  2  separate  stages  in  charge of
  different  jobs. In stage 1, we can use any auxiliary text classification model or tool to provide
  top-K  most  related class labels. These selected class labels will later be used to sample sentence pairs of both positive
  samples  and negative samples for stage 2. Then, in stage 2, we will use sentence pair similarity model to identify the class label with strongest semantic similarity to the user input sentence. Suppose we have a training dataset
  $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$  of  $N$ data points, in which $x_{i}$
  is the user input query and $y_{i}$ is a single class label from a class label
  set  of  $C$  classes  in total. If we want to generate sentence pair dataset
  $\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$,   where   $(x,   x')_{j}$  is
  sentence  pair  grouped  by  2  user  input  queries and $l_{j}\in\{0, 1\}$ is
  similarity class label, see Equ. \ref{eq:similarity label},
  \begin{align}
    l_{j}=
    \begin{cases} \label{eq:similarity label}
      0 & y_{x}\neq y_{x'}\\
      1 & y_{x}=y_{x'}
    \end{cases}
  \end{align}
  directly  from  original dataset $\mathcal{D}$ without any sampling strategy,
  then  $\mathcal{D'}$  will  have  $M=N^2$ data points of sentence pairs, which
  might be too time-consuming for the training process of stage 2 since we often
  have  hundreds  of class labels under task-specific chatbot scenario. Although
  we  may  only  have  a  few data points for a single class label, $M$ is still
  quite a big number.

  Therefore,  we adopt the idea of adversarial negative sampling strategy raised
  by   \cite{bamler2020extreme}.  The  key  idea  of this sampling strategy is to
  train  with  negative  samples that are hard to be distinguished from positive
  samples. Now suppose we have a user input query $x_{i}$ and top-K most related
  class   labels   set  $\mathcal{C'}=\{c'_{1},  \dots,  c'_{K}\}$  provided  by
  auxiliary  model  or  tool in stage 1, where $K$ is a hyperparameter set by us
  that  controls  candidate  class  number,  we  apply  the adversarial negative
  sampling  strategy  into our sentence pair sampling process by the following 2
  steps:

  \textbf{1.  positive  sentence pair sampling}: During the training process, we
  have  to first make sure that the ground truth class label $y_{i}$ for $x_{i}$
  is  within  the  candidate  class  label  set  $\mathcal{C'}$. If $y_{i}\notin
  \mathcal{C'}$,  we  will manually add $y_{i}$ into $\mathcal{C'}$ by replacing
  $c'_{K}$  with  $y_{i}$, since $c'_{K}$ is the least promising candidate label
  according  to  the  auxiliary  model  in stage 1. Afterwards, we will randomly
  sample  $P$  sentences  $\{x'_{0},x'_{1}, \dots, x'_{P}\}$ with the same class
  label as $y_{i}$ from dataset $\mathcal{D}$ to form the set of sentence pairs
  with   positive   label  $\mathcal{P'}_{i}=\{[(x_{i},  x'_{0}),  1],  [(x_{i},
  x'_{1}),   1],   \dots,   [(x_{i},   x'_{P}),  1]\}$,  where  $P$  is  also  a
  hyperparameter  set  by  us  that  controls  the number of sentences we should
  sample from each class.

  \textbf{2.  negative  sentence pair sampling}: As for negative sentence pairs,
  we  will  also  randomly  sample  $P$ sentences for each class in the negative
  candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The class labels
  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set of most confusing class
  labels comparing to the ground truth ${y_{i}}$, so we assume that the sentence
  pairs  with  negative  label grouped by user input query $x_{i}$ and sentences
  with class labels in $\mathcal{C'}\backslash \{y_{i}\}$ are strong adversarial
  negative  samples for sentence-pair similarity model that can help enhance the
  training  speed  and  performance.  According  to  the same method as positive
  sentence  pair sampling, we can obtain the set of sentence pairs with negative
  label  as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{0}), 0], [(x_{i}, x'_{1}), 0], ...,
  [(x_{i}, x'_{P\cdot (K-1)}), 0]\}$

  In  this  way,  we  can generate the sentence pair dataset $\mathcal{D'}$ for
  stage  2 based on the top-K class label set $\mathcal{C'}$ provided by stage 1
  as   $\mathcal{D'}=\{\mathcal{P'}_{i}\cup   \mathcal{N'}_{i}\}_{i=1}^{N}$,  in
  which we will have $M=N\cdot K\cdot P$ data points of sentence pairs in total.

  After  explaining the data sampling principle for sentence pairs and how the 2
  stages  in  SFC  system  connect  with  each other, let's go deeper into model
  structure within these two stages respectively to see how it works.

  \subsubsection*{
    stage 1: classification model for providing top-K candidate class labels
  }

  As for  stage  1,  our  two-stage SFC  system  can adapt to any auxiliary model or even
  searching  tool  based  on  Term Frequency-Inverse Document Frequency (TF-IDF)
  retrieval.  In  this  paper,  however,  we  choose to fine-tune pretrained model such as RoBERTa on our
  task-specific  dataset $\mathcal{D}$ by adding a softmax classifier on top of
  it for better
  overall  performance.  According  to  our  experimental  result,  RoBERTa  can
  guarantee  that  approximately  95\%  ground  truth labels be included in top-5
  candidate  class  labels  with  highest  probability if it has been fine-tuned on
  task-specific  dataset and can obtain a well performance measured by F-score over
  0.75 on this dataset.  So  we use RoBERTa as the encoder of classification model in our setting at
  stage 1 to ensure low limitation for the performance of stage 2.

  For single-sentence text classification tasks, given a data point $x_{i}$ with
  class  label  $y_{i}$  from  dataset  $\mathcal{D}$, RoBERTa takes the final
  hidden  state  $h_{i}$  of the first token [CLS] as the representation for the
  whole   sequence  of  $x_{i}$.  Let's  suppose  we  have  a  linear  layer  as
  ${\Phi}^C_{i}=W^Ch_{i}+b^C$,  where  $W^C$ and $b^C$ are trainable parameters,
  the    probability    score    for    $x_{i}$    can    be    calculated    as
  $p^C_{i}=softmax(W^Ch_{i}+b^C)=softmax({\Phi}^C_{i})$.   Then,  the  loss  for
  stage 1 classification model is shown in Equ. \ref{eq:classification loss}, in
  which  the  first  term  $\varPhi_{i,y_{i}}^C$ encourages high probability for
  prediction of the correct label $y_{i}$.

  \begin{equation}
    \begin{aligned}
      \mathcal{L}^{C}&=\frac{1}{N}\sum_{i=1}^{N}-log(\frac{exp(\varPhi_{i,y_{i}}^C)}{\sum_{j=1}^{C}exp(\varPhi_{i,j}^C)}) \\&=\frac{1}{N}\sum_{i=1}^{N}-\varPhi_{i,y_{i}}^C+log(\sum_{j=1}^{C}exp(\varPhi_{i,j}^C))
      \label{eq:classification loss}
    \end{aligned}
  \end{equation}

  \subsubsection*{
    stage 2: sentence-pair model for multi-task learning
  }

  As for  stage 2, we also choose RoBERTa as the encoder for the sentence pair similarity model. Before sending sentence pairs to Roberta, we
  pack  sentence  pairs  into a single sequence and separate them with a special
  token  [SEP].  Roberta  will  also  add  a  learned  embedding  to every token
  indicating  whether  it belongs to the first sentence or the second one. Then,
  we  utilize  the  cross-attention  mechanism  inside RoBERTa, and then extract
  semantic relationship between 2 sentences from the output softmax layer, which
  is placed on top of the last hidden representation of the [CLS] token.

  Before we start to fine-tune our sentence-pair model on task-specific dataset,
  we   first  fine-tune  RoBERTa  on  Quora  dataset \cite{iyer2017first},  which
  contains  404,290  potential  duplicate question pairs, for transfer learning.
  Then   we   continue  to  do  multi-task  training  on  sentence-pair  dataset
  $\mathcal{D'}$ sampled from the top-K candidate class labels provided by stage
  1.

  The 2 tasks for stage 2 training process are listed below.

  $\bullet$ \textbf{Task 1: Regular sentence-pair similarity task}

  For  sentence  pair semantic similarity task, given a data point $(x, x')_{i}$
  with       similarity       label       $l_{i}$       from       data      set
  $\mathcal{D'}=\{[(x,x')_{i},l_{i}]\}_{i=1}^{M}$,   Roberta   takes  the  final
  hidden  state  $h_{i}$  of the first token [CLS] as the representation for the
  sequence of packed sentence pair $(x, x')_{i}$. Let's suppose we have a linear
  layer  as  ${\Phi}^S_{i}=W^Sh_{i}+b^S$,  where  $W^S$  and $b^S$ are trainable
  parameters,  and  the probability score for $(x, x')_{i}$ can be calculated as
  $p^S_{i}=softmax({\Phi}^S_{i})$.

  Due  to  the  fact that sentence pairs within the same class is much less than
  sentence  pairs  from  different  classes,  the  dataset  $\mathcal{D'}$  is
  imbalanced.  Therefore, we will accommodate a weight variable $w^S = [K-1, 1]$
  to  the  loss  to eliminate the bias brought by imbalanced data, shown in Equ.
  \ref{eq:similarity loss}.

  \begin{equation}
    \begin{aligned}
      \mathcal{L}^{S}&=\frac{1}{M}\sum_{i=1}^{M}-w^S_{l_i}\cdot log(\frac{exp(\varPhi_{i,l_{i}}^S)}{\sum_{j=0}^{1}exp(\varPhi_{i,j}^S)}) \\&=\frac{1}{M}\sum_{i=1}^{M}-w^S_{l_i}\cdot \varPhi_{i,l_{i}}^S+w^S_{l_i}\cdot log(\sum_{j=0}^{1}exp(\varPhi_{i,j}^S))
      \label{eq:similarity loss}
    \end{aligned}
  \end{equation}

  $\bullet$ \textbf{Task 2: Classification task on top-K classes based on sentence-pair model}

  As  for top-K classification task, it shares the same sentence pair similarity
  model   structure  with  task  1.  The  only  difference  is  that  we  add  a
  task-specific  average  pooling  layer,  shown  in  Fig.  \ref{fig:framework} to
  implement the classification task based on sentence-pair model.

  We  already  know  that the size of ${\Phi}^S$ in task 1 is $M\times 2=(N\cdot
  K\cdot  P)\times  2$, where $N$ is the total number of data points in original
  training  data $\mathcal{D}$, $K$ is a hyperparameter that controls the number
  of   candidate   class  labels  provided  by  stage  1,  and  $P$  is  also  a
  hyperparameter that controls the number of sentences we should randomly sample
  from each candidate class labels. Now, for the average pooling layer, we first
  reshape  ${\Phi}^S$  into  the  size  of $N\times K\times P\times 2$, and then
  split  it  into  ${\Phi}^{K,pos}$  and ${\Phi}^{K,neg}$, and both of them will
  have  the  size  of  $N\times  K\times  P$.  In this way, ${\Phi}^{K,pos}$ can
  represent  the  level  of  similarity  for each sentence pair, and in the mean
  time,  ${\Phi}^{K,neg}$  can  represent  the  level  of dissimilarity for each
  sentence  pair. Then, we can do average pooling for each candidate class among
  the  top-K  classes  as  shown  in  Equ.  \ref{eq:average  pooling:1} and Equ.
  \ref{eq:average pooling:2}.

  \begin{align}
    {\xi}_{i,j}^{K,pos} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,pos} \label{eq:average pooling:1}\\
    {\xi}_{i,j}^{K,neg} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,neg}
    \label{eq:average pooling:2}
  \end{align}

  With  the  average pooling result ${\xi}^{K,pos}$ and ${\xi}^{K,neg}$, we also
  generate   a   top-K   class   label   $l^{K}_i\in   [1,\dots,K]$   for   each
  ${\xi}^{K,pos}_{i}$ and ${\xi}^{K,neg}_{i}$. The loss for top-K classification
  task is shown in Equ. \ref{eq:top-k classification loss}. In Equ. \ref{eq: pos
  top-k   loss}   and   Equ.   \ref{eq:   neg   top-k   loss},  the  first  term
  $\xi_{i,l^{K}_{i}}^{K,pos}$  and  $\xi_{i,l^{K}_{i}}^{K,neg}$  encourage  high
  level  of  similarity  and  low  level  of dissimilarity for prediction of the
  correct label $l^{K}_i$ within the top-K candidate classes.

  \begin{align}
    \mathcal{L}^{K} = \frac{1}{N}\sum_{i=1}^{N}(\mathcal{L}^{K,pos}_{i} - \mathcal{L}^{K,neg}_{i})
    \label{eq:top-k classification loss}
  \end{align}
  where,

  \begin{equation}
    \begin{aligned}
      \mathcal{L}^{K,pos}_{i} &= -log(\frac{exp(\xi_{i,l^{K}_{i}}^{K,pos})}{\sum_{j=1}^{K}exp(\xi_{i,j}^{K,pos})}) \\
      &= -\xi_{i,l^{K}_{i}}^{K,pos} + log(\sum_{j=1}^{K}exp(\xi_{i,j}^{K,pos}))
      \label{eq: pos top-k loss}
    \end{aligned}
  \end{equation}

  \begin{equation}
    \begin{aligned}
      \mathcal{L}^{K,neg}_{i} &= -log(\frac{exp(\xi_{i,l^{K}_{i}}^{K,neg})}{\sum_{j=1}^{K}exp(\xi_{i,j}^{K,neg})}) \\
      &= -\xi_{i,l^{K}_{i}}^{K,neg} + log(\sum_{j=1}^{K}exp(\xi_{i,j}^{K,neg}))
      \label{eq: neg top-k loss}
    \end{aligned}
  \end{equation}

  Therefore,  the  overall  loss  function for multi-task learning in stage 2 is
  shown in Equ. \ref{eq: two-stage SFC loss}. The training objective of stage 2 is
  to  minimize  the  weighted  sum  of task-specific losses. Here $\alpha_S$ and
  $\alpha_K$  are  weights  of  task  1  and  task  2  respectively,  which  are
  $\alpha_S=0.125$ and $\alpha_K=0.875$ in our setting. The reason for giving task 2 relatively more weights is that according to the ablation experiments we did in Section \ref{sec:exp}, task 2 contributes more to the overall performance.

  \begin{align}
    \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K
    \label{eq: two-stage SFC loss}
  \end{align}

  \begin{table*}
    \begin{centering}
      \scalebox{0.9}{
      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        \textbf{Dataset} & \textbf{Domain} & \textbf{\#Class} & \textbf{\#Samples} & \textbf{\#Samples/Class} & \textbf{Experimental Settings}\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 1942 & 8 & class sampling\tabularnewline
        \hline 
        HWU64 & intention detection & 64 & [320, 640, 960, 1280, 1920, 3200] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        CLINC150 & intention detection & 150 & [750, 1500, 2250, 3000, 4500, 7500] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        BANKING77 & intention detection & 77 & [385, 770, 1155, 1540, 2310, 3850] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
      \end{tabular}
      }
    \par
    \end{centering}
    \caption{Key statistics for experimental datasets and few shot settings}

    \label{tbe:dataset statistic}
  \end{table*}

  \subsection{joint SFC with joint model structure}
  Despite  the  outstanding experimental result of two-stage SFC on various tasks,
  we still find a limitation brought by the two-stage system structure. In two-stage
  SFC, Stage 1 and stage 2 are separate, so they do not interact with each other
  during  training.  In  this  case,  the  performance  of stage 1 can limit the
  potential  of  stage  2, meanwhile, stage 2 also cannot give training feedback
  back to stage 1 for fine-tuning.

  Therefore,  to  further  improve the overall performance of SFC, we proposed a
  joint  model  structure,  shown  in Fig. \ref{fig:framework}. In
  joint SFC, classification model is being placed in lower layer level to
  dynamically  provide  top-K  candidate  class  labels  for  the  sentence-pair
  similarity  model  placed in higher layer level through a top-K pooling layer.
  In  this way, classification model and similarity model can be merged into one
  single  joint-model  for  multi-task training with sentence pairs sampled from
  varying  candidate  class  labels,  thus  avoiding the limitation brought by 2
  separate stage structure.

  There  are  3  tasks in total during the training process of joint SFC,
  shown  in  Fig. \ref{fig:framework}, and the overall loss is also the weighted
  sum  of  all  the task-specific losses, shown in Equ. \ref{eq: hierachical SFC
  loss}. Here, $\alpha_S$, $\alpha_K$ and $\alpha_C$ are the weights for task 1:
  sentence-pair  similarity,  task  2:  top-K classification and task 3: general
  single   sentence  classification  respectively,  which  are  $\alpha_S=0.25$,
  $\alpha_K=0.25$ and $\alpha_C=0.5$ in our setting.

  \begin{align}
    \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K + \alpha_C \mathcal{L}^C
    \label{eq: hierachical SFC loss}
  \end{align}

  Finally,  when  doing inference, for each user input query $x$, we will choose
  the  class label that contains the sentence $x'$ with highest similarity score
  $\varPhi$,  after  being packed into a sentence pair $(x, x')$ with query $x$,
  as our final result for prediction.

  \section{Experiments}
  \label{sec:exp}

  In  this  section,  we  conduct  extensive  experiments  over  5 task-specific
  datasets  to  evaluate  the  performance  of  our  proposed SFC system and the
  effectiveness  of  the  methodology  of  multi-task  learning  and
  joint-model structure under low-data settings.

  \subsection{Datasets}
  The  datasets  used in our experiments aims to help us study low-resource text
  classification  problem  under  task-specific conversational chatbot scenario.
  Therefore,  we did experiments on datasets under low-resource chatbot corpus setting,
  which  is a large number of class labels (50-250) with a few examples(5-20) in
  short  text  for  each  class, by sampling smaller subsets from the full scale
  when the raw datasets are oversized, see Table \ref{tbe:dataset statistic}. Besides, we also did extensive experiments with larger scale datasets with 30-50 examples per class to test the stability of our proposed SFC system under normal-size data settings, also see Table \ref{tbe:dataset statistic}.

  According to the task and domain, we categorize our experimental datasets into
  the following 3 categories:

  $\bullet$ \textbf{Full-scale real-world task-specific faq chatbot dataset}

  \textbf{ITG}  is  a  real-world  FAQ  chatbot dataset composed of question and
  answer pairs about online English teaching. It contains 3,938 sample questions
  for  228  class labels, where each class label corresponds to an unique answer
  for a bunch of similar sample questions.

  We  divide the whole dataset into 3 folds randomly with approximately the same
  amount  of data within each fold, and then choose 2 folds as the training set,
  the  remaing  1  fold  as  the  test  set for our experiments. We can obtain 3
  different  combinations  among  the  3  folds,  and  we  will  use the average
  experimental  results  of  these 3 combinations as the final evaluation result
  for this dataset.

  $\bullet$ \textbf{Subset of customer review dataset for extreme classification}

  \textbf{Amazon-670K}   is   a   customer   product  review  dataset  for  text
  classification  task from the extreme classification repository with full data
  containing  670,091  class  labels,  285,176  training samples and 150875 test
  samples \cite{bhatia2016extreme}.  As  the  dataset  contains  multiple class
  labels  per  training sample, we keep only the first class label for each data
  point. Besides, we also randomly sampled 250 class labels from all the classes
  that  contain only 5-15 data points. In this way, we simulate the low-resource
  chatbot  corpus setting with the subset of 1,942 training samples and 716 test
  samples for 250 class labels.

  $\bullet$ \textbf{Subset of recently-released intention detection datasets}.

  From our point of view, task-specific chatbot usually starts with small amount
  of  data at the initial stage. As more and more customers start to communicate
  with  the  chatbot,  more data can be extracted from conversation log. Base on
  these  assumptions,  we  do  experiments  on  the following 3 intention detection
  dataset    under    sampling    method    similar    to    the   approach   in
  \cite{casanueva2020efficient}  but  in  more  sophisticated few-shot settings,
  while  keeping the same test sets for each experiment to see the change of the
  improvement scale with decreasing data amount.

  \begin{table*}
    \begin{centering}
      \scalebox{0.55}{
        \begin{tabular}{|c|cccccc|cccccc|cccccc|c|c|}
          \hline 
          \multicolumn{1}{|c|}{\multirow{2}*{\textbf{Models}} }& \multicolumn{6}{c|}{\textbf{CLINC150}}& \multicolumn{6}{c|}{\textbf{BANKING77}}& \multicolumn{6}{c|}{\textbf{HWU64}}& \ \ \ \textbf{ITG} \ \ \ & \textbf{Amazon-670k}\tabularnewline
          \cline{2-21}
          \multicolumn{1}{|c|}{} & $\textbf{5}$ & $\textbf{10}$ & $\textbf{15}$ & $\textbf{20}$ & $\textbf{30}$ & $\textbf{50}$ & $\textbf{5}$ & $\textbf{10}$ & $\textbf{15}$ & $\textbf{20}$ & $\textbf{30}$ & $\textbf{50}$ & $\textbf{5}$ & $\textbf{10}$ & $\textbf{15}$ & $\textbf{20}$ & $\textbf{30}$ & $\textbf{50}$ & \textbf{Full} & \textbf{Full}\tabularnewline
          \hline
          TextCNN & \multirow{2}*{$0.5318$} & \multirow{2}*{$0.6963$} & \multirow{2}*{$0.7609$} & \multirow{2}*{$0.8142$} & \multirow{2}*{$0.8526$} & \multirow{2}*{$0.8867$} & \multirow{2}*{$0.4408$} & \multirow{2}*{$0.6436$} & \multirow{2}*{$0.7366$} & \multirow{2}*{$0.7918$} & \multirow{2}*{$0.8228$} & \multirow{2}*{$0.8619$} & \multirow{2}*{$0.3112$} & \multirow{2}*{$0.4007$} & \multirow{2}*{$0.4823$} & \multirow{2}*{$0.5272$} & \multirow{2}*{$0.5782$} & \multirow{2}*{$0.6262$} & \multirow{2}*{$0.6624$} & \multirow{2}*{$0.4401$}\tabularnewline
          (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          LEAM & \multirow{2}*{$0.7514$} & \multirow{2}*{$0.8203$} & \multirow{2}*{$0.8612$} & \multirow{2}*{$0.8802$} & \multirow{2}*{$0.9010$} & \multirow{2}*{$0.9180$} & \multirow{2}*{$0.5422$} & \multirow{2}*{$0.7812$} & \multirow{2}*{$0.8129$} & \multirow{2}*{$0.8280$} & \multirow{2}*{$0.8610$} & \multirow{2}*{$0.8727$} & \multirow{2}*{$0.4545$} & \multirow{2}*{$0.5554$} & \multirow{2}*{$0.5936$} & \multirow{2}*{$0.6599$} & \multirow{2}*{$0.6855$} & \multirow{2}*{$0.7046$} & \multirow{2}*{$0.7086$} & \multirow{2}*{$0.6091$}\tabularnewline
          (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          BERT-large & \multirow{2}*{$0.8080$} & \multirow{2}*{$0.8904$} & \multirow{2}*{$0.9265$} & \multirow{2}*{$0.9334$} & \multirow{2}*{$0.9497$} & \multirow{2}*{$0.9595$} & \multirow{2}*{$0.5780$} & \multirow{2}*{$0.8004$} & \multirow{2}*{$0.8518$} & \multirow{2}*{$0.8827$} & \multirow{2}*{$0.8858$} & \multirow{2}*{$0.8982$} & \multirow{2}*{$0.4711$} & \multirow{2}*{$0.5963$} & \multirow{2}*{$0.6342$} & \multirow{2}*{$0.7010$} & \multirow{2}*{$0.7117$} & \multirow{2}*{$0.7424$} & \multirow{2}*{$0.7485$} & \multirow{2}*{$0.6658$}\tabularnewline
          (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          ALBERT-xxlarge & \multirow{2}*{$0.8497$} & \multirow{2}*{$0.9008$} & \multirow{2}*{$0.9296$} & \multirow{2}*{$0.9297$} & \multirow{2}*{$0.9466$} & \multirow{2}*{$0.9578$} & \multirow{2}*{$0.5549$} & \multirow{2}*{$0.7981$} & \multirow{2}*{$0.8231$} & \multirow{2}*{$0.8530$} & \multirow{2}*{$0.8571$} & \multirow{2}*{$0.9096$} & \multirow{2}*{$0.4879$} & \multirow{2}*{$0.6116$} & \multirow{2}*{$0.6135$} & \multirow{2}*{$0.6996$} & \multirow{2}*{$0.7094$} & \multirow{2}*{$0.7376$} & \multirow{2}*{$0.7253$} & \multirow{2}*{$0.6893$}\tabularnewline
          (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          RoBERTa-base & \multirow{2}*{$0.8732$} & \multirow{2}*{$0.9254$} & \multirow{2}*{$0.9363$} & \multirow{2}*{$0.9482$} & \multirow{2}*{$0.9558$} & \multirow{2}*{$0.9637$} & \multirow{2}*{$0.7305$} & \multirow{2}*{$0.8654$} & \multirow{2}*{$0.8808$} & \multirow{2}*{$0.9080$} & \multirow{2}*{$0.9061$} & \multirow{2}*{$0.9293$} & \multirow{2}*{$0.5831$} & \multirow{2}*{$0.6790$} & \multirow{2}*{$0.7064$} & \multirow{2}*{$0.7100$} & \multirow{2}*{$0.7320$} & \multirow{2}*{$0.7472$} & \multirow{2}*{$0.7734$} & \multirow{2}*{$0.6708$}\tabularnewline
          (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          RoBERTa-large & \multirow{2}*{$0.8974$} & \multirow{2}*{$0.9372$} & \multirow{2}*{$0.9508$} & \multirow{2}*{$0.9584$} & \multirow{2}*{$0.9621$} & \multirow{2}*{$0.9733$} & \multirow{2}*{$0.7690$} & \multirow{2}*{$0.8728$} & \multirow{2}*{$0.8966$} & \multirow{2}*{$0.9099$} & \multirow{2}*{$0.9227$} & \multirow{2}*{$0.9313$} & \multirow{2}*{$0.6044$} & \multirow{2}*{$0.7002$} & \multirow{2}*{$0.7129$} & \multirow{2}*{$0.7436$} & \multirow{2}*{$0.7493$} & \multirow{2}*{$0.7678$} & \multirow{2}*{$0.7990$} & \multirow{2}*{$0.7156$}\tabularnewline
          (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          RoBERTa-large & \multirow{2}*{$0.8266$} & \multirow{2}*{$0.8861$} & \multirow{2}*{$0.9023$} & \multirow{2}*{$0.9084$} & \multirow{2}*{$0.9090$} & \multirow{2}*{$0.9407$} & \multirow{2}*{$0.757$} & \multirow{2}*{$0.8476$} & \multirow{2}*{$0.8614$} & \multirow{2}*{$0.8743$} & \multirow{2}*{$0.8749$} & \multirow{2}*{$0.8980$} & \multirow{2}*{$0.5425$} & \multirow{2}*{$0.6164$} & \multirow{2}*{$0.6503$} & \multirow{2}*{$0.6729$} & \multirow{2}*{$0.6947$} & \multirow{2}*{$0.7231$} & \multirow{2}*{$0.7418$} & \multirow{2}*{$0.6362$}\tabularnewline
          (similarity) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          2-stage SFC & \multirow{2}*{$0.8979$} & \multirow{2}*{$0.9457$} & \multirow{2}*{$0.9517$} & \multirow{2}*{$0.9591$} & \multirow{2}*{$0.9610$} & \multirow{2}*{$0.9664$} & \multirow{2}*{$0.7975$} & \multirow{2}*{$0.8818$} & \multirow{2}*{$0.8962$} & \multirow{2}*{$0.9109$} & \multirow{2}*{$0.9187$} & \multirow{2}*{$0.9198$} & \multirow{2}*{$0.6477$} & \multirow{2}*{$0.7055$} & \multirow{2}*{$0.7200$} & \multirow{2}*{$0.7232$} & \multirow{2}*{$0.7484$} & \multirow{2}*{$0.7653$} & \multirow{2}*{$0.7972$} & \multirow{2}*{$0.7189$}\tabularnewline
          (task1 only) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          2-stage SFC & \multirow{2}*{$0.9162$} & \multirow{2}*{$0.9424$} & \multirow{2}*{$0.9530$} & \multirow{2}*{$0.9617$} & \multirow{2}*{$0.9633$} & \multirow{2}*{$0.9690$} & \multirow{2}*{$0.7997$} & \multirow{2}*{$0.8823$} & \multirow{2}*{$0.8945$} & \multirow{2}*{$0.9123$} & \multirow{2}*{$0.9236$} & \multirow{2}*{$0.9317$} & \multirow{2}*{$0.6498$} & \multirow{2}*{$0.6980$} & \multirow{2}*{$0.7202$} & \multirow{2}*{$0.7358$} & \multirow{2}*{$0.7498$} & \multirow{2}*{$0.7657$} & \multirow{2}*{$0.8020$} & \multirow{2}*{$0.7311$}\tabularnewline
          (task2 only) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          2-stage SFC & \multirow{2}*{$0.9167$} & \multirow{2}*{$0.9456$} & \multirow{2}*{$0.9571$} & \multirow{2}*{$0.9638$} & \multirow{2}*{$0.9658$} & \multirow{2}*{$0.9753$} & \multirow{2}*{$0.8135$} & \multirow{2}*{$0.8854$} & \multirow{2}*{$0.8931$} & \multirow{2}*{$0.9192$} & \multirow{2}*{$0.9257$} & \multirow{2}*{$0.9339$} & \multirow{2}*{$0.6525$} & \multirow{2}*{$0.7092$} & \multirow{2}*{$0.7168$} & \multirow{2}*{$0.7476$} & \multirow{2}*{$0.7519$} & \multirow{2}*{$0.7696$} & \multirow{2}*{$\textbf{0.8124}$} & \multirow{2}*{$0.7364$}\tabularnewline
          (task1 + task2) & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
          \multirow{2}*{Joint SFC} & \multirow{2}*{$\textbf{0.9231}$} & \multirow{2}*{$\textbf{0.9560}$} & \multirow{2}*{$\textbf{0.9644}$} & \multirow{2}*{$\textbf{0.9669}$} & \multirow{2}*{$\textbf{0.9712}$} & \multirow{2}*{$\textbf{0.9821}$} & \multirow{2}*{$\textbf{0.8270}$} & \multirow{2}*{$\textbf{0.9069}$} & \multirow{2}*{$\textbf{0.9103}$} & \multirow{2}*{$\textbf{0.9209}$} & \multirow{2}*{$\textbf{0.9323}$} & \multirow{2}*{$\textbf{0.9463}$} & \multirow{2}*{$\textbf{0.6697}$} & \multirow{2}*{$\textbf{0.7211}$} & \multirow{2}*{$\textbf{0.7254}$} & \multirow{2}*{$\textbf{0.7497}$} & \multirow{2}*{$\textbf{0.7593}$} & \multirow{2}*{$\textbf{0.7772}$} & \multirow{2}*{$0.8114$} & \multirow{2}*{$\textbf{0.7445}$}\tabularnewline
          & & & & & & & & & & & & & & & & & & & &\tabularnewline
          \hline
        \end{tabular}
      }
      \par
    \end{centering}
    \caption{
        F-scores on five task-specific dataset for text classification in chatbot under low resource. For ITG, we keep the full dataset. For Amazon-670k, we randomly sampled 250 classes with training sample numbers within 5-15 samples per class. For CLINC150, BANKING77, HWU64, we set up various few-shot settings (5/10/15/20/30/50 samples per class) while keeping the test set to be fixed. The highest scores for each data setting are marked in bold.
    }
    \label{tbe:table2}
  \end{table*}

  \textbf{HWU64}   is  a  intention  detection  dataset  designed  for  home  robot
  scenario \cite{liu2019benchmarking}.  It aims at the specific task of capturing
  the  intention  for  different  user  requests  to  home  robot  and  finding the
  corresponding answer. The raw dataset contains 25,716 data points for 64 class
  labels  through  crowdsourcing,  which  raises  much  more  data  than  common
  real-world  circumstances. Therefore, we experiment with the setups where only
  5,  10,  15, 20, 30 and 50 training data points are available for each class label by
  random sampling.

  \textbf{CLINC150}  is a dataset designed for task-oriented systems with 23,700
  queries  that  are  short  and unstructured for 150 intents, in the same style
  made  by real users through crowdsourcing \cite{larson2019evaluation}. Here, we
  sample 5, 10, 15, 20, 30 and 50 training data points for each intention class label by random
  sampling for our experimental setup.

  \textbf{BANKING77}  is  a intention detection dataset for bank customer services.
  The    raw    dataset    contains    13,084   data   points   for   77   class
  labels \cite{casanueva2020efficient}.  Here,  we  sample 5, 10, 15, 20, 30, and 50 training
  data  points  for  each  class  label  by random sampling for our experimental
  setup.

  \subsection{Baselines}
  We choose 3 groups of representative baselines:

  \textbf{Group  1:  single-task  classification  model.} This group of baseline
  models  are  based  on the network structure for classification model.   As   for   non-pretrained   neural   network,  we  choose
  TextCNN \cite{kim2014convolutional}  as  a  representative baseline. As for the
  pretrained  model,  we choose RoBERTa, BERT and ALBERT, all of which are among
  the  most  powerful pretrained models that achieve outstanding results in many
  NLP  tasks,  in  our  experimental  setting. Since we apply RoBERTa in our SFC
  system,   we   naturally  adopt  single-task  classification  model  based  on
  RoBERTa-large  and RoBERTa-base as our baseline models. Besides, we also adopt
  the       recently       released       baseline       pre-trained       model
  ALBERT-xxlarge \cite{lan2019albert}  for  comparison.  Since  we  use few-shot   settings   different from    that of \cite{casanueva2020efficient}  for  those
  intention
  detection  datasets,  we also use BERT-large tuned on our few-shot settings as
  an essential reference baseline for experimental results.

  \textbf{Group  2: single-task similarity model.} This group of baseline models
  are   based   on   network   structure   for   sentence pair similarity   model. We choose RoBERTa-large for the pretrained model part as an
  powerful baseline. But during the evaluation step, we use elastic search to do
  a  rough  filter  for  candidate  sentence  pairs  to  guarantee  a reasonable
  inference time for each user input query.

  \textbf{Group  3:  SFC  system with ablation study.} This group of baseline is
  based  on  the network structure of two-stage SFC. We
  conduct  ablation study by introducing single task SFC system with only Task 1
  or  Task  2  to  investigate  the influence of each component within the total
  loss.

  \subsection{Implementation Details}
  For  our  SFC  system,  we  use RoBERTa-large as the pretrained model for both
  classification  task  and  sentence-pair similarity task. We further fine-tune
  our  SFC  system  on  task-specific  datasets with max sequence length of 128,
  learning  rate  of 1.5e-5 for RoBERTa-large pretrained model and learning rate
  of  5e-4  for  task-specific  layer  with polynomial decay and Adam optimizer.
  Also,  we  set  the  max number of the epoch to 100 and save the best model on
  validation  set based on test result in F-score. We also report the F-score as
  the main evaluation measure for all experimental results.

  For  the  experiments  in  Table  \ref{tbe:table2},  we  empirically  set  the
  hyperparameter  to  be  $K=5$  and  $P=10$  for  all  SFC  variants.  In Table
  \ref{tbe:table3},  we  conduct extensive experiments to see the performance of
  SFC system under various hyperparameter settings.

  \begin{table*}
    \begin{centering}

      \begin{tabular}{|c|c|c|c|c|c|c|}
        \hline 
        \multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
        \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{ITG}}& two-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & joint SFC & $0.7986$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7918$\tabularnewline
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& two-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & joint SFC & $0.7334$ & $0.7445$ & $0.7516$ & $0.7344$ & $0.7373$\tabularnewline
        \hline
      \end{tabular}
      \par
    \end{centering}
    \caption{
      We  show the performance of SFC from different settings of hyperparameters
      for  $K$,  the candidate class number, and $P$, the sample number for each
      class.
    }

    \label{tbe:table3}
  \end{table*}

  \subsection{Result Analysis}

  Table  \ref{tbe:table2}  presents  the  results for each experimental data and
  model  set-up  in detail. The performance of each model is measured in F-score
  on the fixed test set for each dataset. The overall performance of SFC system
  is  better  than  all  the  competitive  single-model  baseline models on five
  experimental   datasets.  This  observation  gives  us  a  general  idea  that
  classification model and similarity model can actually provide each other with
  complementary information, and combining them together into a joint system can
  effectively  enhance the overall capability based on richer extracted semantic
  information.  Furthermore, we also studied  the  relative performance among all the variants of SFC on  different data settings to explore the benefits brought by multi-task training and the joint-model structure.
  With all the experimental results shown in Table \ref{tbe:table2},  we did the following more detailed analysis from several different
  aspect.

  \textbf{Comparison  among  SFC  variants.} First, we turn to compare among the
  four  SFC  variants shown in Table \ref{tbe:table2} to analyze the improvement
  brought  by  multi-task  and  joint-model  structure. We conduct
  ablation  experiments  to study the benefits brought by multi-task training by
  comparing multi-task two-stage SFC with single-task two-stage SFC. We can see from
  Table  \ref{tbe:table2}  that  two-stage  SFC  with  only  task 1 (sentence-pair
  similarity task) degrades by 0.80 percent on average and two-stage SFC with only
  task  2  (top-K  classification  task) degrades by 0.45 percent on average. The
  degradation  indicates  that  training  with these 2 tasks together in SFC can
  effectively  help  the  sentence-pair  model  extract  more knowledge from the
  limited  amount  of data. Besides, task 2 degrades 0.35 percent less than task
  1.  This observation suggests that task 2 may play a relatively more important
  role  in multi-task learning, which leads us to put larger weight $\alpha_{K}$
  to  the  loss  $\mathcal{L}^{K}$  of  task  2  when calculating the total loss
  $\mathcal{L}$.  Also, we can see that the joint SFC steadily outperforms
  two-stage  SFC  on  5  diverse  datasets  by  0.95  percents  on  average.  This
  observation  supports  the idea that the joint model structure can
  alleviate the limitation brought by the lack of interaction between similarity
  model  and  classification  model in two-stage SFC. Therefore, for the following
  analysis,  we  will  focus  on  the comparison between joint SFC and the
  other baseline models.

  \textbf{Analysis  under  few-shot  scenarios.}  Next,  we  analyze the overall
  improvement trend of SFC under various few-shot settings. Our main focus is to
  study chatbot building in common situation during real-world application where
  only  a  few  sample  sentences are available for each class. The experimental
  results  for  the  3  diverse  intention detection datasets (ClINC150, BANKING77,
  HWU64) under few-shot settings(5-20 samples per class) indicate that our proposed joint SFC can
  achieve  2.47  percents  average  improvement  over  one  of the most powerful
  baseline  models,  which  is  RoBERTa-large classification model. Moreover, the improvement in F-score becomes more and more
  prominent  with the decrease of data sample number per class. Our joint
  SFC   can  achieve  4.97  percents  average  improvements  over  RoBERTa-large
  classification model in the most extreme few-shot setting with only 5 training
  samples  per class. This observation indicates that SFC is especially well-performed
  when  being  applied during the initial stage of building task-specific chatbot
  system,  where  the  amount  of  data sample for each class label is extremely scarce. Furthermore, our proposed joint SFC can also achieve 2.04 percents average improvement over RoBERTa-large classification model on all the data settings including the ones with large scale (30-50 samples per class). This experimental result indicates that joint SFC can still steadily outperforms single-task pretrained model baseline even if the data size grows bigger. But the improvement is still more prominent when the data is scarce, which makes SFC an excellent model for low-resource chatbot scenario.

  \textbf{Comparison  between  label  embedding  and  sentence  pair embedding.}
  Afterwards,  we  start  to  compare the performance of SFC with the text-label
  joint  embedding  baseline  named as LEAM. As shown in Table \ref{tbe:table2},
  joint  SFC outperforms LEAM by 11.90 percents on average F-score over 5
  datasets.  Besides,  although the labels within the 3 diverse intention detection
  datasets  (CLINC150,  BANKING77,  HWU64)  are  all  being  well-defined as the
  semantic  summary  of  each intention, joint SFC still outperforms LEAM by
  10.71  percents  on  average  F-score.  This  observation  indicates that it is
  usually  quite  hard  to  use  a  fixed  label  with  only  a few words as the
  representative  of  all  the  sentences  within a class since there are always
  plenty of ways for people to express questions or queries about the same topic
  during   a   conversation  with  chatbot.  Therefore,  learning  the  semantic
  similarity  between  user  query  and  sample  sentences through sentence pair
  embedding  can be more reliable than making decision according to the semantic
  relationship between user query and class labels through label embedding.

  \textbf{Fusion  of  classification  and  similarity  model.} Then, we begin to
  compare  the  general  performance  of SFC with single-task classification and
  similarity  baselines.  As  for  the  classification  model,  joint SFC
  outperforms    TextCNN    by   21.52   percents,   ALBERT-xxlarge   by   7.99
  percents\footnote{We found that the performance of ALBERT.xxlarge is extremely
  unstable  for  classification  task.  Here, we did multiple run with different
  settings  of parameters such as learning rate and batch size for each dataset,
  and  record  the  best  result  we  can  obtain  in  Table \ref{tbe:table2}.},
  BERT-large  by  7.73 percents, RoBERTa-base by 3.79 percents, RoBERTa-large by
  2.04 percents on average F-score over 5 datasets. In this way, we can see that
  RoBERTa  is  the  best  baseline  pre-trained  model for the low-resource text
  classification  task  we  are  studying.  As for the similarity model, we only
  choose  the  strongest  pretrained  model  RoBERTa-large  as our baseline, and
  joint  SFC  achieves  the  improvement  of  7.09  percents  on average.
  Moreover,  joint  SFC's  result on 2 datasets without data sampling for
  few-shot  settings,  which  are  ITG  and  Amazon-670k, are also 1.24 and 2.89
  percents  higher  than  Roberta-large  classification  model, and are 6.96 and
  10.77  percents higher than Roberta-large similarity model. These improvements
  illustrate that putting classification model and similarity model at different
  network  layers  to  cooperate  with  each  other  is  a  successful design of
  joint network structure for joint model, which enables these two models
  assist each other for better overall results.

  \textbf{Settings  of  hyperparameters.} Finally, we analyze the effect brought
  by  different  settings  of  hyperparameters,  see  Table \ref{tbe:table3}. We
  conduct  experiments on 5 different settings of hyperparameters, which are the
  number  of candidate class number $K$ and the number of sentence pairs $P$ for
  each  class, on ITG and Amazon-670K datasets to find the most suitable setting
  of  these  2  hyperparameters.  We  control the total number of sentence pairs
  sampled  for  each  user  query  into  about  the same level of amount(50-60).
  Through  analyzing the experimenal results shown in Table \ref{tbe:table3}, we
  find  that the best setting of hyperparameters might be different for
  datasets of different data size. The class label number of ITG and Amazon-670k
  are  similar  in  our  experimental  setting, which is around 250. However the
  total  amount  of  data  points  of  ITG  is approximately 2 times larger than
  Amazon-670k. Therefore, the setting of $K=5$ and $P=10$ is good enough for ITG
  since  most  of  the  classes  contains  over  10  data  points.  However, for
  Amazon-670k,  the amount of data sample for each class is scarce, thus, we may
  need  more  candidate  classes  to  sample enough effective sentence pairs for
  sentence-pair model layer of SFC. In this way, the setting of $K=10$ and $P=5$
  becomes a good choice. However, in spite of the effect on evaluation performance brought by different settings of hyperparameters, joint SFC still steadily outperforms two-stage SFC by 0.41 percents in overall average F-score on ITG and Amazon-670k datasets.

  In  conclusion,  to  the best of our knowledge, these experimental results place joint SFC as the
  best   performing   model   on   5  diverse  experimental  datasets  for  text
  classification  task, especially when it's under  low-resource task-specific conversational chatbot
  scenario.

  \section{Conclusion}
  In  this  work, we successfully proposed SFC, a joint system of classification
  model  and  sentence-pair similarity model with multi-task learning, which can
  effectively  alleviate  the  limitation of classification model and similarity
  model   for  low-resource  text  classification  tasks.  Experimental  results
  illustrate that our proposed system can achieve great improvement on 5 diverse
  datasets for task-specific chatbot system over several competitive single-task
  baseline  models. Furthermore, we especially analyze the excellent performance
  of SFC when facing the challenge brought by few-shot scenarios, which is quite
  common in real-world application of chatbot system.

  \bibliographystyle{aaai21}
  \bibliography{refs}

\end{document}