### 2020-09-10 chatbot_classification/aaai_2021/abstract.tex
Reorganized.   modified:   paper.tex
  new file:   abstract.tex
  new file:   introduction.tex
  new file:   related_work.tex
  new file:   methodology.tex
  new file:   experiments.tex
  new file:   conclusion.tex

\begin{abstract}
  Task-specific  chatbot systems have gained many important applications, such
  as  Siri,  Google home, and customer service systems, etc. One essential technique 
  underlying these applications  is to detect the real intention behind a user's input.
  The intention detection,  assuming  only  one  intention  in  a  user's input here, can be
  regarded  as  a  short  text  classification problem.  In  practice,  the  number  of
  intentions  can  be hundreds of thousands, as opposed to a few number in the
  traditional text classification problem; creating enough training data from tons of 
  raw user logs with miscellaneous information is not trivial, 
  and generally only 10 to 20 samples per intention are available in  the  early  stage.  
  Thus,  how  to  train  an  effective short-text classification model with
  limited labeled data  in such a practical  and important application is critical. In this work, we propose a
  novel  model, called similarity model fused with classification model (SFC),
  which  combines  a  classification  model  and  a similarity model, and also
  borrows  the  power  of  multi-task  training  as well. We conduct extensive
  experiments  on  4  public  and 1 private datasets. Experimental results
  show that our new model outperforms very strong baselines (i.e., BERT, RoBERTa and
  ALBERT based pretrained models) by over 2 percentages in average F1 score.

\end{abstract}