### 2020-09-10 chatbot_classification/aaai_2021/conclusion.tex

\section{Experiments}
\label{sec:conclusion}

In  this  work,  we  successfully proposed SFC, a joint system of classification
model  and  sentence-pair  similarity  model with multi-task learning, which can
effectively  alleviate  the  limitation  of  classification model and similarity
model   for  low-  resource  text  classification  tasks.  Experimental  results
illustrate  that  our proposed system can achieve great improvement on 5 diverse
datasets  for  task-specific chatbot system over several competitive single-task
baseline models. Furthermore, we especially analyze the excellent performance of
SFC  when  facing  the  challenge  brought by few-shot scenarios, which is quite
common in real-world application of chatbot system.