### 2020-09-10 chatbot_classification/aaai_2021/methodology.tex

\section{Methodology}
In  this section, we describe our proposed system, a similarity model fused with
a classification  model (SFC). We propose two approaches to implement this idea,
two-stage SFC, and its more compact version joint SFC.

\subsection{Two-stage SFC}

Two-stage   SFC  is  composed  with  two  independent stages  in  charge of
different  jobs.

\subsubsection*{
  Stage 1: classification model for providing top-K candidate class labels
} 
We  can  use any auxiliary text classification model or tool,
such as Term Frequency-Inverse Document Frequency (TF-IDF)retrieval, to
provide  top-K  most  related  class  labels. These selected class labels will
later  be  used to sample sentence pairs of both positive samples and negative
samples  for  stage  2.

In  this  work,  we  choose pretrained models to fine tune on our data.
Due to the excellent performance of RoBERTa in this task, compared to other pretrained
models, we use RoBERTa as the encoder of classification model in our setting.

Given   a   data   point   $x_{i}$  with  class  label  $y_{i}$  from  dataset
$\mathcal{D}$, we take the final hidden state $h_{i}$ of the first token [CLS]
encoded  by  RoBERTa  as the representation for the whole sequence of $x_{i}$.
Then  a linear layer is followed to output probabilistic distribution of class
labels,  $softmax(W^Ch_{i}+b^C)=softmax({\Phi}^C_{i})$,  where $W^C$ and $b^C$
are  trainable  parameters.  Then, the loss in stage 1 classification model is
shown in Equ. \ref{eq:classification_loss},

\begin{equation}
  \begin{aligned}
    \mathcal{L}^{C}&=\frac{1}{N}\sum_{i=1}^{N}-log(\frac{exp(\varPhi_{i,y_{i}}^C)}{\sum_{j=1}^{C}exp(\varPhi_{i,j}^C)}) \\
    \label{eq:classification_loss}
  \end{aligned}
\end{equation}

\subsubsection*{
  Stage 2: sentence-pair similarity model based classification with multi-task
  learning
}
We  continue choosing RoBERTa as the main module in this stage that identifies
the class label with strongest semantic similarity to the user input sentence.

Suppose  we have a training dataset $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$ of
$N$  data  points,  in  which $x_{i}$ is the user input query and $y_{i}$ is a
single  class label from a class label set of $C$ classes in total. We 
generate            a sentence            pair            dataset
$\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$ from $\mathcal{D}$,   where   $(x,   x')_{j}$  is
a pair of two history queries \footnote{For simplicity, we use query
to denote a user's question or a response.}, and $l_{j}\in\{0, 1\}$ denotes the
similarity label that only the ones from the same class label are considered
similar. The size of $\mathcal{D'}$  would be up to $O(N^2)$ if no any
sampling strategy is used.  This is quite time-consuming  for the  training
in stage 2 because $N$ can be several hundreds or thousands even under 
few shot setting in task-specific chatbot scenario.

Therefore,  we  adopt  the idea of adversarial negative sampling strategy from
the  work  \cite{bamler2020extreme}.  The  key  idea is to train with negative
samples  that  are hard to be distinguished from positive samples. Now suppose
we  have  a  user  input query $x_{i}$ and top-K most related class labels set
$\mathcal{C'}=\{c'_{1},  \dots,  c'_{K}\}$ provided by auxiliary model or tool
in  stage  1, where $K$ is a hyperparameter to control candidate class number,
we  apply  the  adversarial  negative sampling strategy into our sentence pair
sampling process by the following two steps:

The  first step is \emph{positive sentence pair sampling}. During the training
process,  we  first make sure the ground truth class label $y_{i}$ for $x_{i}$
is  within  the  candidate  class  label  set  $\mathcal{C'}$. If $y_{i}\notin
\mathcal{C'}$,  we  manually  add  $y_{i}$  into  $\mathcal{C'}$  by replacing
$c'_{K}$  with  $y_{i}$, since $c'_{K}$ is the least promising candidate label
according  to  the  auxiliary  model  in stage 1. Afterwards, we will randomly
sample  $P$ sentences $\{x'_{1}, \dots, x'_{P}\}$ with the same class label as
$y_{i}$  from  dataset  $\mathcal{D}$  to  form the set of sentence pairs with
positive  label  $\mathcal{P'}_{i}=\{[(x_{i},  x'_{1}),  1],  \dots,  [(x_{i},
x'_{P}),  1]\}$,  where $P$ is also a hyperparameter set to control the number
of sentences we should sample from each class.

\begin{table*}
  \begin{centering}
    \scalebox{0.9}{
      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        \textbf{Dataset} & \textbf{Domain} & \textbf{\#Class} &
        \textbf{\#Training Samples} & \textbf{\#Samples/Class} & \textbf{Experimental Settings}\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 1942 & 8 & 3-fold\tabularnewline
        \hline 
        HWU64 & intention detection & 64 & [320, 640, 960, 1280, 1920, 3200] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        CLINC150 & intention detection & 150 & [750, 1500, 2250, 3000, 4500, 7500] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        BANKING77 & intention detection & 77 & [385, 770, 1155, 1540, 2310, 3850] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
      \end{tabular}
    }
    \par
  \end{centering}
  \caption{Statistics for all datasets and few shot settings.}

  \label{tbe:dataset statistic}
\end{table*}

The  second  step  is  \emph{negative sentence pair sampling}: As for negative
sentence  pairs,  we will also randomly sample $P$ sentences for each class in
the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The
class  labels  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set  of most
confusing  class  labels comparing to the ground truth ${y_{i}}$, so we assume
that  the  sentence  pairs  with  negative  label  grouped by user input query
$x_{i}$  and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$
are  strong  adversarial  negative  samples for sentence-pair similarity model
that  can  help  enhance  the training speed and performance. According to the
same  method  as  positive  sentence  pair  sampling, we can obtain the set of
sentence  pairs  with  negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{1}),
0], ..., [(x_{i}, x'_{P\cdot K}), 0]\}$

In  this way, we generate the sentence pair dataset $\mathcal{D'}$ for stage 2
based  on  the  top-K  class  label  set $\mathcal{C'}$ provided by stage 1 as
$\mathcal{D'}=\{\mathcal{P'}_{i}\cup  \mathcal{N'}_{i}\}_{i=1}^{N}$,  in which
we have $M=N\cdot K\cdot P$ data points of sentence pairs in total.

Before starting to fine-tune our sentence-pair model on task-specific dataset,
we  first  fine-tune  RoBERTa  on  Quora  dataset  \cite{iyer2017first}, which
contains  404,290  potential  duplicate question pairs, for transfer learning.
This  is  also  the  merit of similarity based model, as labeled task-specific
classification  data  is  hard  to  acquire,  yet  general sentence pairs with
similar semantics can be much easier to acquire.

\subsubsection*{Multi-task based training}
We   continue  to  do  multi-task  training  on  sentence-pair  dataset
$\mathcal{D'}$ sampled from the top-K candidate class labels provided by stage
1. In stage 2, we have two tasks tuning our system.

The first task is \emph{regular sentence-pair similarity task}.
For  sentence  pair semantic similarity task, given a data point $(x, x')_{i}$
with       similarity       label       $l_{i}$       from       data      set
$\mathcal{D'}=\{[(x,x')_{i},l_{i}]\}_{i=1}^{M}$,   Roberta   takes  the  final
hidden  state  $h_{i}$  of the first token [CLS] as the representation for the
sequence of packed sentence pair $(x, x')_{i}$. Let's suppose we have a linear
layer  as  ${\Phi}^S_{i}=W^Sh_{i}+b^S$,  where  $W^S$  and $b^S$ are trainable
parameters,  and  the probability score for $(x, x')_{i}$ can be calculated as
$p^S_{i}=softmax({\Phi}^S_{i})$.

Due  to  the  fact that sentence pairs within the same class is much fewer than
that from  different  classes,  the  dataset  $\mathcal{D'}$  is
quite imbalanced.  Therefore, we will accommodate a weight variable $w^S = [K-1, 1]$
to  the  loss  to eliminate the bias brought by data imbalance, shown in Equ.
\ref{eq:similarity_loss}.

\begin{equation}
  \begin{aligned}
    \mathcal{L}^{S}&=\frac{1}{M}\sum_{i=1}^{M}-w^S_{l_i}\cdot log(\frac{exp(\varPhi_{i,l_{i}}^S)}{\sum_{j=0}^{1}exp(\varPhi_{i,j}^S)}) \\
    \label{eq:similarity_loss}
  \end{aligned}
\end{equation}

The  second  task  is  \emph{classification  on  top-K classes}. As minimizing
similarity  loss  is  not  our  final goal in the intention classification, we
bring  back  classification  loss again. In this task, the network is based on
sentence  pair  similarity  modules.  The  only  difference  is  that we add a
task-specific  average  pooling  layer,  shown  in Fig. \ref{fig:framework} to
accomplish the classification task based on sentence-pair model.

We  already  know  that the size of ${\Phi}^S$ in task 1 is $M\times 2=(N\cdot
K\cdot  P)\times  2$, where $N$ is the total number of data points in original
training  data $\mathcal{D}$, $K$ is a hyperparameter that controls the number
of   candidate   class  labels  provided  by  stage  1,  and  $P$  is  also  a
hyperparameter that controls the number of sentences we should randomly sample
from each candidate class labels. Now, for the average pooling layer, we first
reshape  ${\Phi}^S$  into  the  size  of $N\times K\times P\times 2$, and then
split  it  into  ${\Phi}^{K,pos}$  and ${\Phi}^{K,neg}$, and both of them will
have  the  size  of  $N\times  K\times  P$.  In this way, ${\Phi}^{K,pos}$ can
represent  the  level  of  similarity  for each sentence pair, and in the mean
time,  ${\Phi}^{K,neg}$  can  represent  the  level  of dissimilarity for each
sentence  pair. Then, we can do average pooling for each candidate class among
the  top-K  classes  as  shown  in  Equ.  \ref{eq:average  pooling}.

\begin{align}
  {\xi}_{i,j}^{K,pos} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,pos} \ \ \ \ \ \ \ 
  {\xi}_{i,j}^{K,neg} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,neg}
  \label{eq:average pooling}
\end{align}

With  the  average pooling result ${\xi}^{K,pos}$ and ${\xi}^{K,neg}$, we also
generate   a   top-K   class   label   $l^{K}_i\in   [1,\dots,K]$   for   each
${\xi}^{K,pos}_{i}$ and ${\xi}^{K,neg}_{i}$. The loss for top-K classification
task is shown in Equ. \ref{eq:top-k_classification_loss}. In Equ. \ref{eq: pos
top-k   loss}   and   Equ.   \ref{eq:   neg   top-k   loss},  the  first  term
$\xi_{i,l^{K}_{i}}^{K,pos}$  and  $\xi_{i,l^{K}_{i}}^{K,neg}$  encourage  high
level  of  similarity  and  low  level  of dissimilarity for prediction of the
correct label $l^{K}_i$ within the top-K candidate classes.

\begin{align}
  \mathcal{L}^{K} = \frac{1}{N}\sum_{i=1}^{N}(\mathcal{L}^{K,pos}_{i} - \mathcal{L}^{K,neg}_{i})
  \label{eq:top-k_classification_loss}
\end{align}
where

\begin{equation}
  \begin{aligned}
    \mathcal{L}^{K,pos}_{i} &= -log(\frac{exp(\xi_{i,l^{K}_{i}}^{K,pos})}{\sum_{j=1}^{K}exp(\xi_{i,j}^{K,pos})})
    \label{eq: pos top-k loss}
  \end{aligned}
\end{equation}

\begin{equation}
  \begin{aligned}
    \mathcal{L}^{K,neg}_{i} &= -log(\frac{exp(\xi_{i,l^{K}_{i}}^{K,neg})}{\sum_{j=1}^{K}exp(\xi_{i,j}^{K,neg})}) \\
    \label{eq: neg top-k loss}
  \end{aligned}
\end{equation}

Finally,  the  overall  loss  function for multi-task learning in stage 2 is
shown in Equ. \ref{eq:two-stage_SFC_loss}. The training objective of stage 2 is
to  minimize  the  weighted  sum  of task-specific losses. Here $\alpha_S$ and
$\alpha_K$  are  weights  of  task  1  and  task  2  respectively.

\begin{align}
  \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K
  \label{eq:two-stage_SFC_loss}
\end{align}

\begin{table*}
  \begin{centering}
    \scalebox{0.55}{
      \begin{tabular}{|c|cccccc|cccccc|cccccc|c|c|}
        \hline 
        \multicolumn{1}{|c|}{\multirow{2}*{\textbf{Models}} }& \multicolumn{6}{c|}{\textbf{CLINC150}}& \multicolumn{6}{c|}{\textbf{BANKING77}}& \multicolumn{6}{c|}{\textbf{HWU64}}& \ \ \ \textbf{ITG} \ \ \ & \textbf{Amazon-670k}\tabularnewline
        \cline{2-21}
        \multicolumn{1}{|c|}{} & $\textbf{5}$ & $\textbf{10}$ & $\textbf{15}$
        & $\textbf{20}$ & $\textbf{30}$ & $\textbf{50}$ & $\textbf{5}$ &
        $\textbf{10}$ & $\textbf{15}$ & $\textbf{20}$ & $\textbf{30}$ &
        $\textbf{50}$ & $\textbf{5}$ & $\textbf{10}$ & $\textbf{15}$ &
        $\textbf{20}$ & $\textbf{30}$ & $\textbf{50}$ & \textbf{3-fold} &
        \textbf{3-fold}\tabularnewline
        \hline
        TextCNN & \multirow{2}*{$0.5318$} & \multirow{2}*{$0.6963$} & \multirow{2}*{$0.7609$} & \multirow{2}*{$0.8142$} & \multirow{2}*{$0.8526$} & \multirow{2}*{$0.8867$} & \multirow{2}*{$0.4408$} & \multirow{2}*{$0.6436$} & \multirow{2}*{$0.7366$} & \multirow{2}*{$0.7918$} & \multirow{2}*{$0.8228$} & \multirow{2}*{$0.8619$} & \multirow{2}*{$0.3112$} & \multirow{2}*{$0.4007$} & \multirow{2}*{$0.4823$} & \multirow{2}*{$0.5272$} & \multirow{2}*{$0.5782$} & \multirow{2}*{$0.6262$} & \multirow{2}*{$0.6624$} & \multirow{2}*{$0.4401$}\tabularnewline
        (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        LEAM & \multirow{2}*{$0.7514$} & \multirow{2}*{$0.8203$} & \multirow{2}*{$0.8612$} & \multirow{2}*{$0.8802$} & \multirow{2}*{$0.9010$} & \multirow{2}*{$0.9180$} & \multirow{2}*{$0.5422$} & \multirow{2}*{$0.7812$} & \multirow{2}*{$0.8129$} & \multirow{2}*{$0.8280$} & \multirow{2}*{$0.8610$} & \multirow{2}*{$0.8727$} & \multirow{2}*{$0.4545$} & \multirow{2}*{$0.5554$} & \multirow{2}*{$0.5936$} & \multirow{2}*{$0.6599$} & \multirow{2}*{$0.6855$} & \multirow{2}*{$0.7046$} & \multirow{2}*{$0.7086$} & \multirow{2}*{$0.6091$}\tabularnewline
        (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        \hline
        BERT-large & \multirow{2}*{$0.8080$} & \multirow{2}*{$0.8904$} & \multirow{2}*{$0.9265$} & \multirow{2}*{$0.9334$} & \multirow{2}*{$0.9497$} & \multirow{2}*{$0.9595$} & \multirow{2}*{$0.5780$} & \multirow{2}*{$0.8004$} & \multirow{2}*{$0.8518$} & \multirow{2}*{$0.8827$} & \multirow{2}*{$0.8858$} & \multirow{2}*{$0.8982$} & \multirow{2}*{$0.4711$} & \multirow{2}*{$0.5963$} & \multirow{2}*{$0.6342$} & \multirow{2}*{$0.7010$} & \multirow{2}*{$0.7117$} & \multirow{2}*{$0.7424$} & \multirow{2}*{$0.7485$} & \multirow{2}*{$0.6658$}\tabularnewline
        (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        ALBERT-xxlarge & \multirow{2}*{$0.8497$} & \multirow{2}*{$0.9008$} & \multirow{2}*{$0.9296$} & \multirow{2}*{$0.9297$} & \multirow{2}*{$0.9466$} & \multirow{2}*{$0.9578$} & \multirow{2}*{$0.5549$} & \multirow{2}*{$0.7981$} & \multirow{2}*{$0.8231$} & \multirow{2}*{$0.8530$} & \multirow{2}*{$0.8571$} & \multirow{2}*{$0.9096$} & \multirow{2}*{$0.4879$} & \multirow{2}*{$0.6116$} & \multirow{2}*{$0.6135$} & \multirow{2}*{$0.6996$} & \multirow{2}*{$0.7094$} & \multirow{2}*{$0.7376$} & \multirow{2}*{$0.7253$} & \multirow{2}*{$0.6893$}\tabularnewline
        (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        RoBERTa-base & \multirow{2}*{$0.8732$} & \multirow{2}*{$0.9254$} & \multirow{2}*{$0.9363$} & \multirow{2}*{$0.9482$} & \multirow{2}*{$0.9558$} & \multirow{2}*{$0.9637$} & \multirow{2}*{$0.7305$} & \multirow{2}*{$0.8654$} & \multirow{2}*{$0.8808$} & \multirow{2}*{$0.9080$} & \multirow{2}*{$0.9061$} & \multirow{2}*{$0.9293$} & \multirow{2}*{$0.5831$} & \multirow{2}*{$0.6790$} & \multirow{2}*{$0.7064$} & \multirow{2}*{$0.7100$} & \multirow{2}*{$0.7320$} & \multirow{2}*{$0.7472$} & \multirow{2}*{$0.7734$} & \multirow{2}*{$0.6708$}\tabularnewline
        (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        RoBERTa-large & \multirow{2}*{$\textbf{0.8974}$} &
        \multirow{2}*{$\textbf{0.9372}$} & \multirow{2}*{$\textbf{0.9508}$} &
        \multirow{2}*{$\textbf{0.9584}$} & \multirow{2}*{$\textbf{0.9621}$} & 
        \multirow{2}*{$\textbf{0.9733}$} & \multirow{2}*{$\textbf{0.7690}$} & 
        \multirow{2}*{$\textbf{0.8728}$} & \multirow{2}*{$\textbf{0.8966}$} & 
        \multirow{2}*{$\textbf{0.9099}$} & \multirow{2}*{$\textbf{0.9227}$} & 
        \multirow{2}*{$\textbf{0.9313}$} & \multirow{2}*{$\textbf{0.6044}$} & 
        \multirow{2}*{$\textbf{0.7002}$} & \multirow{2}*{$\textbf{0.7129}$} & 
        \multirow{2}*{$\textbf{0.7436}$} & \multirow{2}*{$\textbf{0.7493}$} & 
        \multirow{2}*{$\textbf{0.7678}$} & \multirow{2}*{$\textbf{0.7990}$} & 
        \multirow{2}*{$\textbf{0.7156}$}\tabularnewline
        (classification) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        \hline
        RoBERTa-large & \multirow{2}*{$0.8266$} & \multirow{2}*{$0.8861$} & \multirow{2}*{$0.9023$} & \multirow{2}*{$0.9084$} & \multirow{2}*{$0.9090$} & \multirow{2}*{$0.9407$} & \multirow{2}*{$0.757$} & \multirow{2}*{$0.8476$} & \multirow{2}*{$0.8614$} & \multirow{2}*{$0.8743$} & \multirow{2}*{$0.8749$} & \multirow{2}*{$0.8980$} & \multirow{2}*{$0.5425$} & \multirow{2}*{$0.6164$} & \multirow{2}*{$0.6503$} & \multirow{2}*{$0.6729$} & \multirow{2}*{$0.6947$} & \multirow{2}*{$0.7231$} & \multirow{2}*{$0.7418$} & \multirow{2}*{$0.6362$}\tabularnewline
        (similarity) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        \hline
        2-stage SFC & \multirow{2}*{$0.8979$} & \multirow{2}*{$0.9457$} & \multirow{2}*{$0.9517$} & \multirow{2}*{$0.9591$} & \multirow{2}*{$0.9610$} & \multirow{2}*{$0.9664$} & \multirow{2}*{$0.7975$} & \multirow{2}*{$0.8818$} & \multirow{2}*{$0.8962$} & \multirow{2}*{$0.9109$} & \multirow{2}*{$0.9187$} & \multirow{2}*{$0.9198$} & \multirow{2}*{$0.6477$} & \multirow{2}*{$0.7055$} & \multirow{2}*{$0.7200$} & \multirow{2}*{$0.7232$} & \multirow{2}*{$0.7484$} & \multirow{2}*{$0.7653$} & \multirow{2}*{$0.7972$} & \multirow{2}*{$0.7189$}\tabularnewline
        (task1) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        2-stage SFC & \multirow{2}*{$0.9162$} & \multirow{2}*{$0.9424$} & \multirow{2}*{$0.9530$} & \multirow{2}*{$0.9617$} & \multirow{2}*{$0.9633$} & \multirow{2}*{$0.9690$} & \multirow{2}*{$0.7997$} & \multirow{2}*{$0.8823$} & \multirow{2}*{$0.8945$} & \multirow{2}*{$0.9123$} & \multirow{2}*{$0.9236$} & \multirow{2}*{$0.9317$} & \multirow{2}*{$0.6498$} & \multirow{2}*{$0.6980$} & \multirow{2}*{$0.7202$} & \multirow{2}*{$0.7358$} & \multirow{2}*{$0.7498$} & \multirow{2}*{$0.7657$} & \multirow{2}*{$0.8020$} & \multirow{2}*{$0.7311$}\tabularnewline
        (task2) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        2-stage SFC & \multirow{2}*{$0.9167$} & \multirow{2}*{$0.9456$} & \multirow{2}*{$0.9571$} & \multirow{2}*{$0.9638$} & \multirow{2}*{$0.9658$} & \multirow{2}*{$0.9753$} & \multirow{2}*{$0.8135$} & \multirow{2}*{$0.8854$} & \multirow{2}*{$0.8931$} & \multirow{2}*{$0.9192$} & \multirow{2}*{$0.9257$} & \multirow{2}*{$0.9339$} & \multirow{2}*{$0.6525$} & \multirow{2}*{$0.7092$} & \multirow{2}*{$0.7168$} & \multirow{2}*{$0.7476$} & \multirow{2}*{$0.7519$} & \multirow{2}*{$0.7696$} & \multirow{2}*{$\textbf{0.8124}$} & \multirow{2}*{$0.7364$}\tabularnewline
        (task1 + task2) & & & & & & & & & & & & & & & & & & & &\tabularnewline
        \hline
        \multirow{2}*{Joint SFC} & \multirow{2}*{$\textbf{0.9231}$} & \multirow{2}*{$\textbf{0.9560}$} & \multirow{2}*{$\textbf{0.9644}$} & \multirow{2}*{$\textbf{0.9669}$} & \multirow{2}*{$\textbf{0.9712}$} & \multirow{2}*{$\textbf{0.9821}$} & \multirow{2}*{$\textbf{0.8270}$} & \multirow{2}*{$\textbf{0.9069}$} & \multirow{2}*{$\textbf{0.9103}$} & \multirow{2}*{$\textbf{0.9209}$} & \multirow{2}*{$\textbf{0.9323}$} & \multirow{2}*{$\textbf{0.9463}$} & \multirow{2}*{$\textbf{0.6697}$} & \multirow{2}*{$\textbf{0.7211}$} & \multirow{2}*{$\textbf{0.7254}$} & \multirow{2}*{$\textbf{0.7497}$} & \multirow{2}*{$\textbf{0.7593}$} & \multirow{2}*{$\textbf{0.7772}$} & \multirow{2}*{$0.8114$} & \multirow{2}*{$\textbf{0.7445}$}\tabularnewline
        & & & & & & & & & & & & & & & & & & & &\tabularnewline
        \hline
      \end{tabular}
    }
    \par
  \end{centering}
  \caption{
    F1 scores  on  five  task-specific  datasets text classification in chatbot
    under low resource. For ITG, we keep the full dataset. For Amazon-670k, we
    randomly  sampled  250  classes  with  training sample numbers within 5-15
    samples  per  class.  For  CLINC150,  BANKING77,  HWU64, we set up various
    few-shot  settings  (5/10/15/20/30/50 samples per class) while keeping the
    test  set to be fixed. The highest scores among all the baseline models and SFC variants for each data setting are both marked
    in bold.
  }
  \label{tbe:table2}
\end{table*}

\subsection{Joint SFC}
In  two-stage  SFC,  Stage  1  and  stage 2 are separate that there is no deep
interaction  with each other during training. In this case, the performance of
stage  1  may  limit  the potential of stage 2; meanwhile, stage 2 also cannot
give  training feedback back to stage 1 for fine-tuning. Therefore, to further
improve  the  overall performance of SFC, we proposed a joint model structure,
shown in Fig. \ref{fig:framework}.

In  joint  SFC,  classification  model is being placed in lower layer level to
dynamically  provide  top-K  candidate  class  labels  for  the  sentence-pair
similarity  model  placed in higher layer level through a top-K pooling layer.
In  this way, classification model and similarity model can be merged into one
single  joint-model  for  multi-task training with sentence pairs sampled from
varying  candidate  class  labels,  thus  avoiding the limitation brought by 2
separate stage structure.

There  are 3 tasks in total during the training process of joint SFC, shown in
Fig. \ref{fig:framework}, and the overall loss is also the weighted sum of all
the   task-specific  losses,  shown  in  Equ.  \ref{eq:joint_SFC_loss}.  Here,
$\alpha_S$,  $\alpha_K$  and  $\alpha_C$  represent  the  weights  for  task 1
sentence-pair  similarity  in  Equ.  \ref{eq:similarity_loss},  task  2  top-K
classification  in Equ. \ref{eq:top-k_classification_loss}, and task 3 general
single       sentence       classification      respectively      in      Equ.
\ref{eq:classification_loss}.

\begin{align}
  \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K + \alpha_C \mathcal{L}^C
  \label{eq:joint_SFC_loss}
\end{align}