### 2020-09-10 chatbot_classification/aaai_2021/related_work.tex

\section{Related work}

Another  common  approach  is  text  classification  neural  network  based on
label-word     joint     embedding,    such    as    LEAM \cite{wang2018joint},
MTLE \cite{zhang2017multi}   and   EXAM \cite{du2019explicit}.   This   approach
involves  label  embedding when constructing the text-sequence representation.
Take  LEAM  as  an  example,  the  text classification model is implemented by
jointly    embedding    the    word    and    label   in   the   same   latent
space \cite{wang2018joint}. Afterwards, the text representation are constructed
based  on  text-label  compatibility through attention mechanism. In this way,
additional  sources  of  semantic  information  from class labels can be fully
leveraged, which is quite helpful especially under few-shot scenario.

As for \emph{label-word joint
embedding  approach},  it  poses  noticeable requirements on label information.
However,  in  our  cases,  since  we have large amount of class labels without
being well-defined for question answering chatbot, the only thing we can do is
to  choose  one  of the user utterance as the class label to represent the key
information  of  the  whole class. In this way, the label will become not only
prolix  but  also  biased,  which  will  bring  some  negative effect to label
embedding.

Another   line   of  interesting  work  is  the  joint  training  of  sentence
classification   and  NER  (todo:  reference).  If  both  the  city  and  date
information  are  recognized,  then  they  are good indicators that this query
might  be  booking  an  ticket. This work can also be combined in our proposed
framework.