### 2020-11-06 chatbot_classification/aaai_2021/rebuttal.txt
renamed:    aaai reviews for rebuttal.pdf -> aaai_reviews_for_rebuttal.pdf   new file:   rebuttal.txt

----------------------------- * For Reviewer #3 * ------------------------------

### Question 1: Was the pre-trained model used in baseline were fine-tuned …

Response:
Thanks  for  your invaluable question. All pre-trained model based baselines are
fine-tuned. Sorry for our careless assumption that in NLP almost all pre-trained
model based systems have to be fine-tuned.

Usually  in our experiments, RoBERTa and BERT model layers are tuned, regardless
of  being  in  baselines or SFC, with a learning rate of 5e-4, and task-specific
layers with a learning rate of 1e-4. This setting is empirically efficient.

Besides, all other settings, like the polynomial decay, Adam optimize, and using
a development set, are the same.

#####  Question  2:  the baseline can be stronger, such as joint-training of NER
and intent detection using BERT.

Response:
Thanks  for such a nice question. We did notice this line of work, introduced in
line 154. We can put the analysis below in our next version. Appreciate.

First,  the  datasets,  usually used for both intent and NER tasks, namely Snips
and ATIS, have quite limited intent number, 7 and 21 respectively. However, they
define  quite  rich  NER  labels,  72  and  120  respectively. In our researched
applications,  the available labeled data is often limited, then NER module in a
joint  model  might not be sufficiently trained. However, for some popular NERs,
address, person, time, city, there are more available data for better training.

Now  considering  a  joint  model,  our  framework  is  still  supportive of NER
training. In the first stage of SFCs, which is a BERT-like classification model,
we  can  trivially develop a fourth objective, as NER. This is a clear effective
idea. Refer to "BERT for Joint Intent Classification and Slot Filling".

However,  the  four  public datasets and one private dataset do not have any NER
information.  Thus,  in  this  work,  we  exclusively compare the classification
performance.

#####  Question  3:  -  There  is  not  much  what  ...  exactly was improved by
introducing the similarity model.

Response: 
Sorry  for  the  unclear explanation. This is really difficult to exactly answer
why.   We   tried  to  explain  this  point  in  "Does  joint  SFC  improve  the
classification  in  stage  1" in the experimental part. Our experiment, table 3,
show  that  SFC  improves  the top-5 accuracy in the first stage, at the cost of
top-1 accuracy, to improve the final performance.

Hope this kind of clarify this issue. Appreciate!

----------------------------- * For Reviewer #4 *  -----------------------------

###  Question  1: "classification model cannot use data augmentation method" and
"similarity  model  doesn't  perform  well  on a new task-specific dataset". Why
putting ... together ...?

Response:
Thanks  for  your  question.  This  is an invaluable opinion that we should have
spent effort to clarify it.

In  the  classification  model,  let  (x, y) be the input feature and label of a
sample.  In  our target application, we need enough labeled data like such pairs
(x,  y).  However, we are lacking of those pairs, except spending lots of effort
labeling them manually. Even though in a second application, we have enough data
like  (x,  y), and because the label definitions of two applications are totally
different,  then  we  have no way to bring the data in the second application to
the first application.

This  is  quite  different  in  the similarity model which only models two x’s
similarity, and this definition is more general. In inference, given a x_new, we
target at searching the most similar x, modeled by similarity(x_new, x), and use
the  label  of  this  x as the label of x_new. Let’s continue with the example
above.  In  the  third application, though its training data is from a different
domain,  yet the similarity model (x1, x2) can be used in the first application,
as the label definitions do not matter here.

I hope this explanation can provide more information for you. Appreciate!

###  Question  2:  How  to solve the error propagation problem?

Response:
Thanks  for  you  in-depth  question.  The  multi-stage  modeling is popular and
inevitable  in  NLP.  Hence  a  general  idea is joint modeling which trains all
stages simultaneously.

Our  two-stage  SFC  does  have  the  issue pointed out by you. Thus, we further
propose  joint  SFC  to  alleviate  this  issues. However, this issue can not be
solved ideally.

Our  table 3 explores the intermediate result in the joint SFC. We can see that,
the  top-5  accuracy  in  the  first  stage  got  improved, at the cost of top-1
accuracy, in exchange of the final improved performance.

I hope this looks a bit clearer for you. Appreciate!

### Question 2: The authors simply putting a classification model and a matching
model together, and the authors haven't made any improvement.

Response: Sorry for confusing you. 
First,  the combination of two rather different models are not trivial, as their
structures  are  hard  to combine efficiently. In comparison, the two losses are
trivial to combine.

In  our practical applications, such as chatbot, we are often facing the problem
of  lacking  sufficient  labeled  data.  The  classification  model  has  a high
accuracy,  while  quite  hard to utilize external data; the similarity is on the
contrary. This is the motivation of combing two models.

I hope this can help you understand better. Appreciate!

----------------------------- * For Reviewer #5 * ------------------------------

###  Question  1: Can you explain the impact of the fine-tuning process on Quora
dataset?

Response:  This  is a really interesting problem. We did notice this issue, yet we
think  this is not influencing the fairness of our model over ALL baselines. Due
to the target workload of experiments, we did not explain much about this issue.
Let's clarify in-depth now.

First the merit of a similarity model is that we can expand our training data.
Due to page limit, could refer to the response to the first question of reviewer
#4?

As  our  researched  problem  is  the  application  without sufficient data, the
similarity  model baseline is also influenced due to insufficient in-domain data
to  a  great extent. Thus, generally the similarity baseline has to use external
data.

Then this merit is naturally inherited in our SFC models. When comparing SFCs to
a  similarity model baseline, our experiments are fair; when comparing SFCs to a
classification, supporting external data is also our advance, which is fair too.
Hence, in order to reduce the explanation workload, we just use this setting.

### Question 2: Will you release you code if your paper were accepted?

Definitely! I am stilling considering how to release the codes without violating
the rule of releasing author information.

----------------------------- * For Reviewer #6 *  -----------------------------

###  Question  1:  The method seems time-consuming, especially the second stage.

Response: thanks for this good question.

Regarding  the  training time: the time of SFCs is actually the summation of the
classification  model  baseline  and  the  similarity  model  baseline.  In  our
settings,  the  classification  model  needs  about  5  hours in a single Nvidia
RTX8000 GPU; the similarity model needs about 12 hours; our model needs about 14
hours due to faster convergence.

Regarding the inference time, the classification model baseline is about 0.2 ms;
the  similarity  model  baseline  is  about  0.5 ms; our joint SFC needs 0.6 ms.
Actually  both  the  similarity  model  baseline  and  our  joint SFC can be GPU
parallelized in the similarity computation, if we need pretty fast response time
in a real product.

As  our  researched problem is the application without sufficient training data,
hence, the extra training cost from the external data might not be a problem.

Thanks for pointing out. We should have included in our paper!