### 2021-05-16 chatbot_classification/03.emnlp_2021/abstract.tex
modified:   abstract.tex   modified:   experiments.tex   modified:   paper.ver00.pdf

\begin{abstract}
  Task-specific chatbot systems have gained many important applications, such as smart  speaker,  customer  service system. 
  One fundamental module behind them is detecting intent of a user's input, and can be modeled as a short text classification problem.  
  However, in  the early stage of building a chatbot, collecting enough labeled data  for  hundreds  of  thousands  of  user  intents is expensive. 
  Popular classification models, direct mapping a query to an intent, have a high precision, while depending on enough task-specific labels information. 
  In comparison, similarity models, modeling  similarity  of two queries instead, can utilize additional out-of-domain data, while  having  a  relatively  lower  precision,  due  to  the  discrepancy  of similarity loss and real classification loss. 
  In this work, we propose a novel model,  called  similarity  model  fused  with  classification model (SFC), to combine  the  merits of the two kinds of models in the framework of multi-task training.  
  Our  extensive  experiments  on  6  public  and  1 private datasets demonstrate  that  our  systems outperform very strong baselines (i.e., RoBERTa based  pretrained  model,  joint model with NER), especially with insufficient data.

\end{abstract}