### 2021-05-16 chatbot_classification/03.emnlp_2021/appendix.tex
modified:   appendix.tex   modified:   experiments.tex   modified:   introduction.tex   modified:   methodology.tex

\appendix

\newpage
\section{Appendix}

\subsection{Related work}
We supplement some work that is very different from classic classification models.

One  line  of  work  is  based  on  label-word  joint  embedding,  such  as LEAM
\cite{wang2018joint}, MTLE \cite{zhang2017multi} and EXAM \cite{du2019explicit}.
It  poses  extral  requirements  on label information. However in our case, many
class  labels  are  close  to  each  other,  and not well-defined. Though we can
present the class label with one of user's utterances from that class, the label
will  become  not  only prolix but also biased.

Another   line   of   interesting   work  is  the  joint  training  of  sentence
classification     and     NER,     such    as    
\cite{kruengkrai2020improving,zhang2020graph,hakkani2016multi,liu2016attention,goo2018slot}.  
For  example,  If  both  the  city and date information are
recognized,  then  they  are good indicators that this query might be booking an
ticket.

\subsection{Settings  of  hyperparameters} 
In  Table \ref{tbe:table3}, we explore the influence from hyperparameters, which
are  the  number  of  candidate  class  labels $K$ in stage 1, and the number of
sampled  sentence  pairs  for  each class $P$ in stage 2, on ITG and Amazon-670K
datasets.  As we control the total number of sampled sentence pairs to about 50 to 60, we
roughly obatin five combinations of hyperparameters.

We find  that the best setting of hyperparameters is different from datasets
of  different  data  size.  The  class label number of ITG and Amazon-670k are
similar  in  our  experimental setting, which is around 250. However the total
amount of data points of ITG is approximately 2 times larger than Amazon-670k.
Therefore,  the  setting of $K=5$ and $P=10$ is good enough for ITG since most
of  the  classes  contains  over 10 data points. However, for Amazon-670k, the
amount  of  data  sample  for  each  class  is  scarce, thus, we may need more
candidate  classes to sample enough effective sentence pairs for sentence-pair
model  layer  of  SFC.  In this way, the setting of $K=10$ and $P=5$ becomes a
good choice.

However,  in  spite of the effect on evaluation performance brought by different
settings  of hyperparameters, joint-SFC still steadily outperforms two-stage SFC
by  0.41  percent  points  in  overall  average  F1 score on ITG and Amazon-670k
datasets.

\subsection{Implementation Details}
For  our  SFC  systems,  we  use RoBERTa-large as the pretrained model for both
classification  task  and  sentence-pair similarity task. We fine-tune
our  SFC  system  on  our datasets with max sequence length of 128,
learning  rate  of 1.5e-5 for RoBERTa-large pretrained model and learning rate
of  5e-4  for  task-specific  layer  with polynomial decay and Adam optimizer.

In two-stage SFC, in equ. \ref{eq:two-stage_SFC_loss} the weights of two tasks
are  set  as  $\alpha_S=0.125$ and $\alpha_K=0.875$ in our setting. The reason
for  giving  task  2 relatively more weights is that according to the ablation
experiments  we  did  in Section \ref{sec:exp}, task 2 contributes more to the
overall  performance.  In  joint-SFC,  in  equ.  \ref{eq:joint_SFC_loss}, the
weights  of  three  tasks  are  set  as  $\alpha_S=0.25$,  $\alpha_K=0.25$ and
$\alpha_C=0.5$  in  our setting. Overall, the task weights are less sensitive
in joint-SFC than in two-stage SFC.

For  the  experiments  in  Table  \ref{tbe:table2},  we  empirically  set  the
hyperparameter  to  be  $K=5$  and  $P=10$  for  all  SFC  variants.  In Table
\ref{tbe:table3},  we  conduct extensive experiments to see the performance of
SFC system under various hyperparameter settings.

As for the training time, in our settings, our joint-SFC model needs about 14
hours to reach convergence using a single Nvidia RTX8000 GPU. In comparison, the classification baseline model needs about 5 hours and the similarity baseline model
needs about 12 hours, which indicates that our joint-SFC model spends less training
time than the summation of two baseline models due to faster convergence.

As for the inference time, the classification model baseline is about 0.2 ms;
the similarity model baseline is about 0.5 ms and our joint-SFC model takes only 0.6 ms.
Actually both the similarity model baseline and our joint-SFC can be GPU
parallelized in the similarity computation, if we need pretty fast response time
in a real product.