### 2021-05-16 chatbot_classification/03.emnlp_2021/experiments.tex

\section{Experiments}
\label{sec:exp}

\subsection{Datasets}
Our experiments aim to study the short text classification in the low-resource environment  popular  in  the  task-specific conversational chatbot application.
The   datasets  adopted  in  our  experiments, showed in \ref{tbe:dataset statistic} \footnote{All settings of six public datasets would be released to github soon.}, typically  contain comparatively  large  number  of  class labels, ranging from several dozens to hundreds,  and  each  class  label  is  associated  with  a handful of labeled queries with each  query  being usually  one  sentence.

\begin{enumerate}
  \item \emph{ITG},  is  a proprietary FAQ dataset from real-world
  chatbot  project,  which is composed of question and answer pairs about online
  English teaching. It contains 3,938 sample questions for 228 class labels, and
  each class label corresponds to a unique answer.

  \item \emph{Amazon-670K}, is a customer product review dataset for
  text  classification  task  from  the  extreme  classification repository. The
  complete  dataset  contains 670,091 class labels, 285,176 training samples and
  150,875   testing   samples   \cite{bhatia2016extreme}.  As  each  sample  may
  correspond  to  multiple  class labels, we keep only the first one. We further
  filter  out the class labels that only associated to training samples within the amount of 5 to 15 to  mimic  the  few-shot chatbot scenario. From them, we sample 250 class
  labels as well as their training samples to form a subset with 2658 samples.

  \item \emph{HWU64}, is an intent detection dataset designed for
  home  robot  scenario \cite{liu2019benchmarking}. It aims at the specific task
  of  capturing  the  intent  for  different  user requests to home robot and
  finding  the corresponding answer. The raw dataset contains 25,716 data points
  for 64 class labels through crowdsourcing.

  \item \emph{CLINC150},  is  a dataset designed for task-oriented
  systems  with  23,700 queries that are short and unstructured for 150 intents,
  in    the    same   style   made   by   real   users   through   crowdsourcing
  \cite{larson2019evaluation}.

  \item \emph{BANKING77}, is an intent detection dataset for bank
  customer  services.  The  raw dataset contains 13,084 data points for 77 class
  labels \cite{casanueva2020efficient}.

  \item \emph{FRAMES}, is a collection of multi-domain dialogues dealing with
hotel bookings. The raw data is consists of 1369 human-human
dialogues with slot filling information. There are 30,522 utterances in total for 21 intents and 49 slots \cite{asri2017frames}.

  \item \emph{ATIS}, is a popular dataset for intent detection of flight reservations with slot filling\cite{tur2010left}. The raw data's training, development and test sets contain 4,478, 500 and 893 utterances, respectively. There are 120 slot labels and 21 intent types for the training set.

\end{enumerate}

Regarding  the first two datasets, we conduct 3-fold cross validation experiments that treat 70   percent   as  training,  15  percent  as  validation  and  testing respectively,  and  report  the  averaged  testing results. 
Regarding the last five datasets, we use a sampling method similar to that from \cite{casanueva2020efficient}   yet  in  a  more  sophisticated  few-shot settings.  
We  fix  a  test  data for each one, and examine their performances using  5,  10,  15,  20,  30,  or 50 samples per class label respectively for training.  
This  is important, as in practice a task-specific chatbot usually starts  with  merely  a  handful of labeled data available in the early stage.
Besides,  in  the  active  learning framework, building an effective auxiliary system  with limited resource is also quite important to help developers label new data more efficiently.

\subsection{Baselines}
Our baselines comprise of four styles of systems.

\begin{enumerate}
  \item   
  \emph{Non-pretrained   model  based  classification  system},  which consists  of  a typical CNN based TextCNN \cite{kim2014convolutional}, and a label-embedding based LEAM \cite{wang2018joint}. 
  Regarding TextCNN, the input is from RoBERTa tokenization, and the kernels are set as 1, 2, 3, 4, 5. 
  Regarding LEAM, a literal class label is required, which is available only in HWU64, CLINC150, BANKING77. 
  Thus, for other two datasets, we have to use its class number instead.  
  Empirically, a  non-pretrained  model  based  system  are  not  performing  as  well as a pretrained  models  based in many NLP tasks.
  Our listing these  non-pretrained model based systems here is serving a comprehensive performance comparison on our datasets.

  \item 
  \emph{Pretrained model based classification system}, which consists of BERT \cite{devlin2018bert},  RoBERTa \cite{liu2019roberta},  and  ALBERT  \cite{lan2019albert}  based.  
  These pretrained models  are  practically  proven  to  achieve  outstanding  performances  in classification task and other NLP tasks.

  \item   
  \emph{Pre-trained   model   based   similarity   model}.  
  As RoBERTa is empirically found more effective than other pretrained models  in  the  short  text  classification  task  in  our  experiments, we  choose RoBERTa-large  based implementation.
  In inference,  we  use  an  elastic search to find a set of potential candidate labels for the similarity model, to guarantee a reasonable running time.

  \item 
  \emph{Pre-trained model based joint text classification and NER model}. 
  We choose the work from \cite{chen2019bert}, which utilizes a RoBERTa-large model to jointly train two losses, as our baseline system.
  Our  SFC framework  is also conveniently supportive of extra NER loss that in  stage  1 a fourth task for NER loss is trivially derived  based on the  output states from a pretrained model.

\end{enumerate}

All our pretrained model   baselines  are  fine-tuned  on  our datasets.
Especially, the similarity  model  baseline  and all similarity modules in SFCs use extra Quora  dataset  \cite{iyer2017first}  to enhance system performance. 
This is also one of merits in SFC as it support adopting out-of-domain data in comparison to a classification based model.

Our SFC implementations include three 2-stage SFCs, which uses task 1 and 2 with ablation in Table \ref{tbe:table2} ; one joint-SFC which is trained jointly on all three tasks in Table \ref{tbe:table2}; one joint-SFC with an additional fourth NER task in Table \ref{tbe:addtional ner experiments}.

All setting details can be found in the appendix due to the space limitation.

\subsection{Result Analysis}
We  report  the F1 score \footnote{In the multi-class and multi-label case, this
is  the  average  of  the F1 score of each class with weighting depending on the
numbers  in  each class.} as the main evaluation measure for all experiments in
Table \ref{tbe:table2} and Table \ref{tbe:addtional ner experiments}.

\subsubsection*{Are multi-task and joint training working?} 
First, we compare the four  SFC  variants shown in Table \ref{tbe:table2} to analyze the improvement
brought  by  multi-task  and  joint-model  structure.

\begin{table}
  \begin{centering}
    \scalebox{0.58}{
      \begin{tabular}{|c|ccc|ccc|}
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{\textbf{Models}} }& \multicolumn{3}{c|}{\textbf{FRAMES}}& \multicolumn{3}{c|}{\textbf{ATIS}}\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & $\textbf{10}$ & $\textbf{20}$ & $\textbf{50}$
        & $\textbf{10}$ & $\textbf{20}$ & $\textbf{50}$\tabularnewline
        \hline
        RoBERTa-large & \multirow{2}*{${0.3456}$} & \multirow{2}*{${0.4043}$} & \multirow{2}*{${0.4262}$} & \multirow{2}*{${0.9349}$} & \multirow{2}*{${0.9757}$} & \multirow{2}*{${0.9800}$}\tabularnewline
        (classification) & & & & & & \tabularnewline
        \hline
        RoBERTa-large & \multirow{3}*{${0.3520}$} & \multirow{3}*{${0.4088}$} & \multirow{3}*{${0.4353}$} & \multirow{3}*{${0.9560}$} & \multirow{3}*{${0.9706}$} & \multirow{3}*{$\textbf{0.9832}$}\tabularnewline
        (classification & & & & & & \tabularnewline
        w/ NER) & & & & & & \tabularnewline
        \hline
        Joint-SFC & \multirow{2}*{${0.3843}$} &
        \multirow{2}*{${0.4130}$} & \multirow{2}*{${0.4390}$} &
        \multirow{2}*{${0.9618}$} & \multirow{2}*{${0.9761}$} & 
        \multirow{2}*{${0.9825}$}\tabularnewline
        (w/o NER) & & & & & & \tabularnewline
        \hline
        Joint-SFC & \multirow{2}*{$\textbf{0.3925}$} &
        \multirow{2}*{$\textbf{0.4420}$} & \multirow{2}*{$\textbf{0.4456}$} &
        \multirow{2}*{$\textbf{0.9639}$} & \multirow{2}*{$\textbf{0.9784}$} & 
        \multirow{2}*{${0.9826}$}\tabularnewline
        (w/ NER) & & & & & & \tabularnewline
        \hline
      \end{tabular}
    }

  \end{centering}
  \caption{
    F1 scores for joint sentence classification and NER training. 
    Regarding both FRAMES and ATIS dataset, we set up various few-shot settings (10/20/50 samples per intent) while fixing the original test set.
  }
  \label{tbe:addtional ner experiments}
\end{table}

\begin{table}
  \begin{centering}
    \scalebox{0.85}{
        \begin{tabular}{|c|c|c|c|}
        \hline
        F1 score & 2-stage SFC & joint-SFC & Gap \tabularnewline
        \hline
        top-1 accuracy  & \textbf{81.50} & 81.07 & -0.47 \tabularnewline
        top-5 accuracy & 94.01 & \textbf{94.30} & +0.29 \tabularnewline
        \hline
      \end{tabular}
    }
    \par
  \end{centering}
  \caption{The average classification accuracy in percentage in stage 1 on all five dataset.}
  \label{tbe:top1_5_accuracy}
\end{table}

Comparing with training 2-stage SFC with multi-task, training with only task
1,  namely the sentence pair similarity model, degrades by 0.8 percent
point  on average, and training with only task 2, namely the top-K based
classification  task,  degrades  by  0.45  percent  point  on  average.  These
degradation  indicates  multi-task training in 2-stage SFC is helpful for the
system  performance.  Besides, task 2 plays a relatively more important role in
multi-task learning, and this aligns with their optimal weight settings.

Joint-SFC consistently outperforms 2-stage SFC with multi-task training  
on 5 diverse datasets by 0.95 percents  on  average. 
This observation supports the idea that the joint model
structure  can  alleviate  the  limitation  brought by the lack of interaction
between  two  stages.  In  following analysis, we will focus on the comparison
between joint-SFC and the other baseline models.

\subsubsection*{Is  the  fusion of classification and similarity models
helpful?}  We  compare  joint-SFC with classification, similarity and NER joint model based baselines respectively.

\begin{table}
  \begin{centering}
  \scalebox{0.58}{
    \begin{tabular}{|c|c|c|c|c|c|c|}
      \hline 
      \multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
      \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
      \hline
      \multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
      \cline{2-7}
      \multicolumn{1}{|c|}{} & joint-SFC & $0.7986$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7918$\tabularnewline
      \hline
      \multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
      \cline{2-7}
      \multicolumn{1}{|c|}{} & joint-SFC & $0.7334$ & $0.7445$ & $0.7516$ & $0.7344$ & $0.7373$\tabularnewline
      \hline
    \end{tabular}
    \par
  }
  \end{centering}
  \caption{
    We  show the performances of SFC from different settings of
    hyperparameters, $K$  denoting the candidate class number from stage 1,
    $P$ denoting the number of sampled sentence pair in stage 2. 
  }
  \label{tbe:table3}
\end{table}

Regarding the  classification  models,  joint-SFC
outperforms RoBERTa-large based, the strongest among all baselines, 
by 2.04 percent points on average F1 score over 5 datasets.  
Especially, ALBERT.xxlarge based is rather unstable  in this scenario. Thus, 
we did multiple run with different settings and  report the  best  ones here. 
Comparatively,  RoBERTa  based is  the  most stable and has the best performance 
n almost all experiment settings.

Regarding  the  similarity  model,  as  analyzed  above,  we  only  choose the
RoBERTa-large based implementation as our baseline. Joint-SFC achieves more improvement of
7.09  percent points  on average F1 score.

As for the intent classification and NER joint model, we choose the state-of-the-art RoBERTa-large model with NER loss as our strongest baseline. For ablation study, we did experiments on joint-SFC model with and without NER loss respectively. From Table \ref{tbe:addtional ner experiments}, we can see that our joint-SFC without NER loss still outperforms RoBERTa with NER loss baseline by 0.86 percent points on average. This indicates that the fusion of classification and similarity can even works better than fusing with additional NER information under few-shot chatbot scenario. Moreover, after adding NER loss to joint-SFC, it achieves further improvement by 1.66 percent points over RoBERTa with NER loss baseline. Especially for FRAMES dataset, which is a much more difficult task in comparison with ATIS, our joint-SFC with NER loss achieves a bigger improvement of 2.8 percent points on average. In this way, we show the compatibility of our joint-SFC model with extra semantic information, as well as the stable outstanding performance of joint-SFC when facing hard tasks with few training data.

The above analysis illustrates that suitable fusing these two kinds of models
can take their advantages.

\subsubsection*{Does joint-SFC improve the classification in stage 1?}
We  compare  the  top-1  prediction and top-5 candidate labels in stage 1 from
2-stage SFC and joint-SFC in Table \ref{tbe:top1_5_accuracy}. We should note
that  2-stage  SFC  does not optimize the stage 1 model, which is a RoBERTa
based classification model.

The joint training does not improve the classification performance in stage 1,
measured  by  top-1  accuracy.  However,  it  improves  the  quality of top-5
candidate  class  labels, which can further improve the effect of sentence pair
similarity training. The reason is that, a classification model
inherently optimizes the objective loss, 0-1 error of top-1 here; the sentence
pair similarity model in the joint-SFC poses more positive effect in optimizing the 
candidate labels.

\subsubsection*{How does training sample size influence SFC?} 
We analyze the overall  improvement  trend  of  joint-SFC under various few-shot
settings, and show results in Fig. \ref{fig:trend}.

Our main focus  is  to  study  chatbot  building  in common situation during real-world
application  where  only  a few sample sentences are available for each class.
The  experimental  results  for  the  3  diverse  intent detection datasets
(ClINC150,  BANKING77,  HWU64)  under  few-shot  settings, 5 to 50 samples per
class,  indicate that our proposed joint-SFC can achieve 2.47 percents average
improvement   over  one  of  the  most  powerful  baseline  models,  which  is
RoBERTa-large  classification  model.  Moreover,  the  improvement  in F1 score
becomes  more  and  more prominent with the decrease of data sample number per
class.

\begin{figure}[t]
  \begin{centering}
    \includegraphics[scale=0.2]{picture/improvement_trend.jpg} 
    \par
  \end{centering}
  \caption{
    Improvements  from  Joint  SFC  over  RoBERTa  based  classification  with
    different size of training data.
  }
  \label{fig:trend}
\end{figure}

Especially,  in the most extreme few-shot setting with only 5 training samples
per  class  our joint-SFC achieves 4.97 percent points improvements on average
over  RoBERTa-large classification model. This observation indicates that SFC
is  especially  well-performed  when being applied during the initial stage of
building  task-specific  chatbot  system,  where the amount of data sample for
each class label is extremely scarce.

Furthermore,  our  proposed  joint  SFC can also achieve 2.04 percents average
improvement  over  RoBERTa-large classification model on all the data settings
including  the  ones  with  large  scale,  30  to  50  samples per class. This
indicates that joint-SFC can still steadily outperforms pretrained model based
baseline even if the data size grows bigger. However, the improvement is still
more prominent when the data is scarce, which makes SFC an excellent model for
low-resource chatbot scenario.

\subsubsection*{How does joint-SFC relate to the label  embedding based  LEAM?}
LEAM performs consistently better than TextCNN by a large margin in all
datasets. TextCNN is practically efficiently in task-specific chat
applications, and LEAM shows its power in modeling labeling information.

LEAM  does not perform as well as SFCs, since the first intuitive reason is it
does  not use pretrained model as encoding module; and another critical reason
is  in  task-specific  chat applications, it is common to have many intents
that are close to some others with a minor difference, and it is hard and even
impossible  to  name  each  intent  with  a short clear name. One candidate
solution  is  setting  a most standard sample as the label of that class. When
using  non-pretrained  models  base  LEAM,  this is applicable. Yet when using
pretrained  model  based, as the all labels can be hundreds of thousands, then
this  can not be accommodated in a poplar 32 G Tesla V100 GPU. Actually, joint
SFC  can  be  kind  of  understood  as  a  generalization form of LEAM. In the
scenario when there is no clear class label, the relationship between a sample
and  a class label is implicitly encoded as that between a sample and another
sample  from  the  same  class, and this turns into a sentence pair similarity
model.