### 2021-05-18 chatbot_classification/03.emnlp_2021/appendix.tex
modified:   appendix.tex   modified:   experiments.tex

\newpage

\appendix

\section{Appendix}
\subsection{Related work}
We supplement some work that is very different from classic classification models.

One  line  of  work  is  based  on  label-word  joint  embedding,  such  as LEAM
\cite{wang2018joint}, MTLE \cite{zhang2017multi} and EXAM \cite{du2019explicit}.
It  poses  extral  requirements  on label information. However in our case, many
class  labels  are  close  to  each  other,  and not well-defined. Though we can
present the class label with one of user's utterances from that class, the label
will  become  not  only prolix but also biased.

Another   line   of   interesting   work  is  the  joint  training  of  sentence
classification     and     NER,     such    as    
\cite{kruengkrai2020improving,zhang2020graph,hakkani2016multi,liu2016attention,goo2018slot}.  
For  example,  If  both  the  city and date information are
recognized,  then  they  are good indicators that this query might be booking an
ticket.

\subsection*{Relation to the label embedding based LEAM}
LEAM shows its superiority over TextCNN in all datasets in our experiments, by modeling labeling information.

LEAM does not perform as well as SFCs, since the first intuitive reason is it does not use a pretrained model as its input encoding module; 
another critical reason is in task-specific chat applications, it is common to have many similar intents, and it becomes hard and even impossible to name each intent with a short clear name to feed into LEAM model.
One candidate solution is setting a most standard sample as the label of the class. 
When using non-pretrained models base LEAM, this is applicable. 
Yet when using a pretrained model based, as all labels can be hundreds of thousands, then the training can not be accommodated in a poplar 32 G Tesla V100 GPU.

Actually, joint SFC can be kind of understood as a generalization form of LEAM. 
When there is no clear class labels, the relationship between a sample and a class label is implicitly encoded as that between a sample and another sample from the same class, and this turns into a sentence pair similarity model.