### 2021-05-18 chatbot_classification/03.emnlp_2021/introduction.tex
modified:   experiments.tex   modified:   introduction.tex   modified:   methodology.tex   new file:   paper.ver07.pdf

\section{Introduction}
\label{sec:intro}

Task-specific conversational chatbot \cite{wen2016network} has been applied into many practical products. 
A popular one is the smart speaker, e.g. Alex, Siri, Google home. 
Another important one is the customer service system, which greatly help human agents handle miscellaneous customer's questions.
No matter these applications involves single- or multi-round conversations, a critical step is to identify the intent behind a user's question or response. 
The detected intent with its associated attributes is then mapped into a predefined dialog logic to obtain a suitable response to return to customers.

In the early stage of building a chatbot , sufficient labeled data is often expensive to obtain to render the system to achieve a strong performance.
People thus have to filter out enough typical users' utterances from tons of real conversation logs, and label them with proper intents. 
In practice, this procedure can be improved by active-learning like operations. 
A basic intent detection system with very limited manually labeled data can be built first, and then it filters out a set of potential data with high confidences for humans to label.
The new data is added to train a better system again. 
This procedure iterates until the system reach a high performance.
Therefore, it is meaningful to address the challenge of short-text classification \cite{sriram2010short, chen2019deep, phan2008learning,yan2009dynamic,hua2015short} problem under the few-shot setting \cite{yu2018diverse}.

Existent approaches for intent detection can be roughly categorized as text classification model and text similarity model.

The first one, \emph{text classification model}, includes a variety of work.
From traditional machine learning models like SVM \cite{suykens1999least}, boosting tree \cite{tu2005probabilistic}, to neural networks \cite{wen2016network}, such as convolutional neural networks (CNNs) \cite{kim2014convolutional,zhang2015character,conneau2016very} and long short term memory networks (LSTMs) \cite{mousa2017contextual,liu2016recurrent}, and then to the most popular pretrained language models based, such as BERT \cite{devlin2018bert}, RoBERTa \cite{liu2019roberta} and etc. 
Especially, pretrained models based \cite{vaswani2017attention} tends to be more helpful in the few-shot scenario \cite{yu2018diverse, madabushi2020cost} to alleviate the dearth of training data.
Another two interesting lines of work, label-word joint models, and joint NER and classification are discussed in the related work of appendix.

The second one, \emph{text similarity model}, is usually employed to calculate how similar between an input text and a historical text in the repository. 
The associated label of the most similar historical text is returned as the label of the text to query \cite{jafarpour2010filter, leuski2011npceditor}. 
A popular and effective methodology is adopting pretrained models to model the similarity calculation as a binary classification problem.

Despite plenty of success of these two methodologies, they still have some limitations in this task especially when data is insufficient.

Regarding the \emph{text classification model}, it learns a function that directly maps an input query into its expected label.
The training requires data in the form of a pair of a query and a task dependent label. 
Yet it is quite hard to adopt existent labeled data from other domains, since their label definitions are often incompatible to each other. 
Though we can use some data augmentation methods to alleviate the paucity of data, e.g., translation based paraphrasing methods, the resulting benefit is still limited because of the data being homogeneous, in the cases that labeled data is rather insufficient.

Regarding the \emph{similarity model}, the foremost advantage is it does not pose any restriction on the label definition of user intents, but aims to learn a function that measure how similar of two sentences are. 
Then in the intent detection task, we are only concerned which labeled query is the most similar to the current input, and then use its label as the output.
This results in a possibility that even though the labeled data in current domain is scarce, we may borrow additional data from another domain to help enhance the similarity model to make a better intent prediction.
Nevertheless, just as every coin has two sides, the high flexibility of similarity model leads to its worse performance compared to a classification model, as it is not trained towards the goal in the intent detection task which is actually a 0-1 loss minimization.
More, it requires an auxiliary model to narrow down candidate class labels to speed up the similarity calculation next step, and this auxiliary model should be of a good balance in precision and recall.

The above limitations motivate us to propose a system that may take both the high performance from a classification model and the ability of supporting out-of-domain data from a similarity model.
We call our system as SFC, short for similarity model fused with classification model, shown in Fig. \ref{fig:framework}. 
In order to train effectively, we further borrow the multi-task learning \cite{caruana1993multitask,collobert2008unified, liu2019multi}.
Our basic idea come naturally, and the first implementation consists of two stages. 
In the first stage, we use an auxiliary model to select top-K most possible labels for an input. 
This model can be an elastic search \cite{divya2013elasticsearch} or a text classification model trained on current domain. 
In the second stage, we build a classification model which is composed of several similarity modules. 
Then, this structure derives two goals to train towards, the task-specific classification loss , and the similarity loss on both in-domain and out-of-domain data. 
In this version, the two stages are independently optimized, so we call it \emph{2-stage SFC}.

We further find that the quality of outputs from the first stage might limit the final performance of the system, since those outputs are fixed in the whole system training procedure. 
This observation motivates us to continue improving the above 2-stage SFC into a joint training setting, and we call this second version as \emph{joint-SFC}. 
In this way, the first stage model can be optimized by the optimization in the second stage so that it can provide better candidates to improve the performance of the second stage in turn.

Since SFC is in the framework of multi-task training, its inherently supports adding additional tasks, such as NER, to gain further improvement.

\begin{figure*}[t]
  \begin{centering}
    \includegraphics[scale=0.66]{picture/picture4} 
    \par
  \end{centering}
  \caption{
    \textbf{Network Structure of SFC:} 2-stage SFC and joint-SFC are sharing the same network, with the only difference whether the two stages being jointly trained.
  }
  \label{fig:framework}
\end{figure*}